### 2023-02-14 bibat/{{cookiecutter.repo_name}}/docs/report.qmd
Switch markdown docs format to quarto

---
title: "{{ cookiecutter.project_name }}"
subtitle: "A report created using bibat version {{ cookiecutter.bibat_version }}"
author: "{{ cookiecutter.author_name }}"
bibliography: bibliography.bib
reference-location: margin
format:
  html: default
  pdf: default
---

Write your report here in [pandoc
markdown](https://pandoc.org/MANUAL.html#pandocs-markdown).

Turn it into a pdf by running `quarto render report.qmd` from the project root.

Refer to entries in `bibliography.bib` using their bibtex keys like
this: @gelmanBayesianWorkflow2020.

Include images like this:

![This is an example image](img/example.png)

Include margin notes like this^[I'm a margin note!]

Write inline code like this:

```{python}
def print_hello():
    print("Hello!")
```

Check out the [quarto guide](https://quarto.org/docs/guide/) for more
information about how to document your work using quarto.

# References

A list of references should appear here: