### 2023-05-04 bibat/{{cookiecutter.repo_name}}/docs/report.qmd

---
title: "{{ cookiecutter.project_name }}"
resource_path: ["../"]
subtitle: "A report created using bibat version {{ cookiecutter.bibat_version }}"
author: "{{ cookiecutter.author_name }}"
bibliography: bibliography.bib
reference-location: margin
format:
  html: default
  pdf: default
---

Write your report here in [pandoc
markdown](https://pandoc.org/MANUAL.html#pandocs-markdown).

Create Pdf and Html files based on this document by running `quarto render
report.qmd` from the project root. If you want to remove some of these formats
or add a new one, you can do so by editing the `format` field of the
document's yaml front matter. For example if you want to create a version that
your collaborators can open in Microsoft Word:

```{.yaml}
---
format:
  html: default
  pdf: default
  docx: default
---
```

Refer to entries in `bibliography.bib` using their bibtex keys like
this: @gelmanBayesianWorkflow2020.

Include images like this:

![This is an example image](img/example.png)

Note that you can also include links to images in the main project's `plots`
folder:

![This image comes from the main analysis!](../plots/posterior_predictive_comparison.png)

Include margin notes like this^[I'm a margin note!]

Write code blocks like this:

```{python}
def print_hello():
    print("Hello!")
print_hello()
```

Note that when quarto rendered the document it evaluated the code above and
printed the results! Quarto supports execution of Python, R and Julia code, and
can print figures that or even interactive widgets. If you want to include a
block of code from one of these languages, but don't want quarto to execute it,
or want to [customise what quarto does with the code in another
way](https://quarto.org/docs/computations/execution-options.html) make sure the
first line is `#| `, e.g.:

```{python}
#| eval: false
def print_hello():
    print("Hello!")
print_hello()
```

Check out the [quarto guide](https://quarto.org/docs/guide/) for more
information about how to document your work using quarto.

# References

A list of references should appear here: