### 2023-06-02 bibat/examples/baseball/docs/report.qmd
update baseball example

---
title: "Modelling at-bats in baseball using the generalised Pareto distribution"
subtitle: "A report created using bibat version 0.1.6"
author: "Teddy Groves"
bibliography: bibliography.bib
reference-location: margin
toc: true
format:
  html:
    embed-resources: true
filters:
  - include-code-files
---

I used to do a lot of statistical analyses of sports data where there was a
latent parameter for the player's ability. You can see an example
[here](https://github.com/teddygroves/cricket).

It was a natural choice to use a Bayesian hierarchical model where the abilities
have a location/scale type distribution with support on the whole real line, and
in fact this worked pretty well! You would see the kind of players at the top
and bottom of the ability scale that you would expect.

Still, there were a few problems. In particular the data were typically
unbalanced because better players tended to produce more data than worse
players. The result of this was that my models would often inappropriately think
the bad players were like the good players: they would not only tend to be too
certain about the abilities of low-data players, but also be biased, thinking
that these players are probably a bit better than they actually are. I never
came up with a good way to solve this problem, despite trying a lot of things!

Even though I don't work with sports data very much any more, the problem still
haunted me, so when I read [this great case
study](https://mc-stan.org/users/documentation/case-studies/gpareto_functions.html#conclusion-on-the-data-analysis)
about geomagnetic storms it gave me an idea for yet another potential solution.

The idea was this: just as data about intense storms that cause electricity
problems tell us about a tail of the bigger solar magnetism distribution, maybe
data about professional sportspeople is best thought about as coming from a tail
of the general sports ability distribution. If so, maybe something like the
[generalised pareto
distribution](https://en.wikipedia.org/wiki/Generalized_Pareto_distribution)
might be better than the bog standard normal distribution for describing the
pros' abilities.

I thought I'd test this out with some sports data, and luckily there is a really
nice baseball example on [the Stan case studies
website](https://mc-stan.org/users/documentation/case-studies/pool-binary-trials.html),
complete with [data from the 2006 Major league
season](https://github.com/stan-dev/example-models/blob/master/knitr/pool-binary-trials/baseball-hits-2006.csv).
After I posted some early results [on the Stan discourse
forum](https://discourse.mc-stan.org/t/baseball-analysis-using-latent-generalised-pareto-distribution/29222),
other users suggested that it might be interesting to model similar data from
different seasons. This data can now be found quite easily on the [baseball
databank](https://baseballdb.lawlesst.net/).

With a few datasets and at least two different statistical models to consider,
the full analysis looked like it would be a little too big to fit in a trivial
file structure, so it seemed like a good opportunity to showcase my
batteries-included Bayesian analysis template package
[bibat](https://github.com/teddygroves/bibat).

The rest of this vignette describes how I used bibat to (relatively) painlessly
see if my generalised Pareto distribution idea would work.

Check out the analysis's [github
repository](https://github.com/teddygroves/baseball) for all the details.

# Setup

First I installed bibat in my global Python 3.11 environment with this command:

```zsh
> pip install bibat
```

Next I ran bibat's wizard like this:

```{.zsh}
> bibat
Welcome to the Batteries-Included Bayesian Analysis Template!
What is your project called? [Project Name]: Baseball
What should the project repository be called [baseball]: baseball
What is your name? [Author name]: Teddy Groves
What is your email? [author@email.com]:
Who should be the code of conduct contact? [author@email.com]:
Please briefly describe your project [A short description of the project.]: Comparison of distributions for modelling baseball hitting
Choose an open source license from these options: (MIT, BSD-3-Clause, No license file) [MIT]:
How would you like to document your project? (Quarto, Sphinx, No docs) [Quarto]:
Would you like to create a tests directory? [y]: n
Would you like to create a .github directory? [y]: n
>
```

After I answered the wizard's questions bibat creted a new folder called
`baseball` that looked like this:

```{.zsh}
baseball
├── CODE_OF_CONDUCT.md
├── LICENSE
├── Makefile
├── README.md
├── bibat_version.txt
├── data
│   └── raw
│       ├── raw_measurements.csv
│       └── readme.md
├── docs
│   ├── bibliography.bib
│   ├── img
│   │   ├── example.png
│   │   └── readme.md
│   ├── report.html
│   ├── report.pdf
│   └── report.qmd
├── baseball
│   ├── __init__.py
│   ├── data_preparation.py
│   ├── fitting_mode.py
│   ├── inference_configuration.py
│   ├── investigate.ipynb
│   ├── notebooks
│   ├── prepare_data.py
│   ├── readme.md
│   ├── sample.py
│   ├── stan
│   │   ├── custom_functions.stan
│   │   ├── multilevel-linear-regression.stan
│   │   └── readme.md
│   ├── stan_input_functions.py
│   └── util.py
├── inferences
│   ├── fake_interaction
│   │   └── config.toml
│   ├── interaction
│   │   └── config.toml
│   └── no_interaction
│       └── config.toml
├── plots
│   ├── posterior_ll_comparison.png
│   └── posterior_predictive_comparison.png
└── pyproject.toml
```

This folder implements bibat's example analysis – a comparison of linear
regression with two different design matrices. To check that everything was
working correctly I ran the analysis like this:

```zsh
$ cd baseball
$ make analysis
```

This ran without errors, running some data preparation functions and then analysing the data with Stan in a few configurations.

# Getting raw data

To fetch raw data from the internet, I wrote a new script in the file
`baseball/fetch_data.py`:

```{.python include="../baseball/fetch_data.py"}
```

To get the files I ran the script:

```zsh
> source .venv/bin/activate
(baseball) > python baseball/fetch_data.py
Fetching 2006 data from https://raw.githubusercontent.com/stan-dev/example-models/master/knitr/pool-binary-trials/baseball-hits-2006.csv
Writing 2006 data to data/raw/2006.csv
Fetching bdb-main data from https://raw.githubusercontent.com/chadwickbureau/baseballdatabank/master/core/Batting.csv
Writing bdb-main data to data/raw/bdb-main.csv
Fetching bdb-post data from https://raw.githubusercontent.com/chadwickbureau/baseballdatabank/master/core/BattingPost.csv
Writing bdb-post data to data/raw/bdb-post.csv
Fetching bdb-apps data from https://raw.githubusercontent.com/chadwickbureau/baseballdatabank/master/core/Appearances.csv
Writing bdb-apps data to data/raw/bdb-apps.csv
```

Finally, I removed the example analysis's raw data:

```zsh
(baseball) > rm data/raw/raw_measurements.csv
```

# Preparing the data

The first step in preparing data is to decide what prepared data looks like for
the purposes of our analysis. Bibat provides dataclasses called `PreparedData`
and `MeasurementsDF` to help get started with this, which I found in the file
`baseball/data_preparation.py`.

As it happens, prepared data looks very similar in this analysis and the
example. All I had to do was change the `MeasurementsDF` definition a
little^[note that this class uses [pandera](https://pandera.readthedocs.io), a
handy library for defining what a pandas dataframe should look like]:

```python
class MeasurementsDF(pa.SchemaModel):
    """A PreparedData should have a measurements dataframe like this.

    Other columns are also allowed!
    """

    player_season: Series[str]
    season: Series[str] = pa.Field(coerce=True)
    n_attempt: Series[int] = pa.Field(ge = 1)
    n_success: Series[int] = pa.Field(ge = 0)
```

The next step is to write functions that return `PreparedData` objects. In this
case I wrote a couple of data preparation functions: `prepare_data_2006` and
`prepare_data_bdb`:

```{.python include="../baseball/data_preparation.py" start-line=103}
```

I had to update the `prepare_data` function to take into account the
two raw data sources:

```{.python include="../baseball/data_preparation.py" start-line=35 end-line=54}
```

To finish off I deleted the unused global variables `NEW_COLNAMES`,
`DROPNA_COLS` and `DIMS`, then checked if the function `load_prepared_data`
needed any changes: I was pretty sure it didn't.

To check that all this worked, I ran the data preparation script manually:^[I
could also have just run `make analysis` again. This would have caused an error
on the step after prepare_data.py, which is fine!]

```{.zsh}
> source .venv/bin/activate
(baseball) > python baseball/prepare_data.py
```

Now the folder `data/prepared/bdb` contained a file
`data/prepared/bdb/measurements.csv` that looked like this:

```{csv}
,player_season,season,n_attempt,n_success
0,abreujo02,2018,553,180
1,acunaro01,2018,487,178
3,adamewi01,2018,322,112
6,adamsla01,2018,29,10
7,adamsma01,2018,337,104
8,adamsma01,2018,277,92
9,adamsma01,2018,60,12
```

# Specifying statistical models

I wanted to test two statistical models: one with the modelling the distribution
of per-player logit-scale at-bat success rates as a normal distribution with
unknown mean and standard deviation, and another where the same logit-scale
rates have a generalised Pareto distribution.

So, given a table of $N$ player profiles, with each player has $y$ successes out
of $K$ at-bats and an unknown latent success rate $\alpha$, I wanted to use this
measurement model:

$$
y \sim \text{binomial logit}(K, \alpha)
$$

In the generalised Pareto model I would give the $\alpha$s this prior model,
with the hyperparameter $\text{min }\alpha$ assumed to be known exactly and $k$
and $\sigma$ given prior distributions that put the $\alpha$s in the generally
plausible range of between roughly 0.1 and 0.4.

\begin{align*}
\alpha &\sim GPareto(\text{min }\alpha, k, \sigma)
\end{align*}

In the normal model I would use a standard hierarchical regression model with an
effect for the log-scale number of at-bats to attempt to explicitly model the
tendency of players with more appearances to be better:

\begin{align*}
\alpha &\sim Normal(\mu + b_{K} \cdot \ln{K}, \tau) \\
\end{align*}

Again I would choose priors for the hyperparameters that put most of the alphas
between 0.1 and 0.4.

To implement these models using Stan I first added the following function to the
file `custom_functions.stan`. This was simply copied from [the relevant part of
the geomagnetic storms case
study](https://mc-stan.org/users/documentation/case-studies/gpareto_functions.html#conclusion-on-the-data-analysis).

```{.stan}
real gpareto_lpdf(vector y, real ymin, real k, real sigma) {
  // generalised Pareto log pdf
  int N = rows(y);
  real inv_k = inv(k);
  if (k<0 && max(y-ymin)/sigma > -inv_k)
    reject("k<0 and max(y-ymin)/sigma > -1/k; found k, sigma =", k, ", ", sigma);
  if (sigma<=0)
    reject("sigma<=0; found sigma =", sigma);
  if (fabs(k) > 1e-15)
    return -(1+inv_k)*sum(log1p((y-ymin) * (k/sigma))) -N*log(sigma);
  else
    return -sum(y-ymin)/sigma -N*log(sigma); // limit k->0
}
```

Next I wrote a file `gpareto.stan`:

```{.stan include="../baseball/stan/gpareto.stan"}
```

Finally I wrote a file `normal.stan`:

```{.stan include="../baseball/stan/normal.stan"}
```

# Generating Stan inputs

Next I needed to tell the analysis how to turn some prepared data into a
dictionary that can be used as input for Stan. Bibat assumes that this task is
handled by functions that live in the file `baseball/stan_input_functions.py`,
each of which takes in a `PreparedData` and returns a Python dictionary. You
can write as many Stan input functions as you like and choose which one to run
for any given inference.

I started by defining some Stan input functions that pass arbitrary prepared
data on to each of the models:^[Note that this code uses the scipy function
`logit`, which it imported like this: `from scipy.special import logit`]

```{.python include="../baseball/stan_input_functions.py" start-line=11 end-line=34}
```

But why stop there? It can also be useful to generate Stan input data with a
model, by running it in simulation mode with hardcoded parameter values. Here
are some functions that do this for both of our models:

```{.python include="../baseball/stan_input_functions.py" start-line=36}
```

# Specifying inferences

Now all the building blocks for making statistical inferences – raw data, data
preparation rules, statistical models and recipes for turning prepared data
into model inputs – were in place. The last step before actually running Stan
was to write down how put the blocks together. Bibat has another concept for
this, called 'inferences'.

An inference in bibat is a folder containing a special file called
`config.toml`. This file sets out what inferences you want to make: which
statistical model, which prepared data function, which Stan input function,
which parameters have which dimensions, which sampling modes to use and how to
configure the sampler. The folder will later be filled up with the results of
performing the specified inferences.

I started by deleting the example inferences and creating two fresh folders,
leaving me with an `inferences` folder looking like this:

```zsh
inferences
├── gpareto2006
│   └── config.toml
└── normal2006
    └── config.toml
```

Here is the file `inferences/gpareto2006/config.toml`:

```{.toml include="../inferences/gpareto2006/config.toml"}
```

Here is the file `inferences/normal2006/config.toml`:

```{.toml include="../inferences/normal2006/config.toml"}
```

Note that:

- The Stan file, prepared data folder and Stan input function are referred to by
  strings. The analysis should raise an error if you enter a non-existing value.
- Both inferences are set to run in "prior" and "posterior" modes – the other
  pre-defined mode is "kfold", but you can also write your own!
- You can enter arbitrary arguments to cmdstanpy's `CmdStanModel.sample` method
  in the `[sample_kwargs]` table.
- You can enter mode-specific overrides in `[mode_options.<MODE>]`. This can be
  handy if you want to run more or fewer iterations for a certain mode.

With all these changes made I ran `make analysis` again. There were no errors
until after sampling, which was expected as I hadn't yet customised the
investigation code, and I saw messages indicating that Stan had run. I also
found that the `inferences` subfolders had been populated:

```zsh
inferences
├── gpareto2006
│   ├── config.toml
│   └── idata.json
└── normal2006
    ├── config.toml
    └── idata.json
```

# Investigating the inferences

Now that the inferences are ready it's time to check them out. Bibat provides a
Jupyter notebook at `baseball/investigate.ipynb` for exactly this purpose. The
notebook's main job is to create plots and save them in the `plots` directory
when it is executed with the command `jupyter execute investigate.ipynb`, which
is the final step in the chain of commands that is triggered by `make
analysis`.

A notebook is arguably a nicer home for code that creates plots than a plain
python script because it allows for literate documentation and an iterative
workflow. A notebook makes it easy to, for example, add some code to change the
scale of a plot, execute the code and see the new results, then update the
relevant documentation all in the same place.

The code from the example analysis's notebook for loading `InferenceData` was
reusable with a few tweaks to avoid missing file errors, so I kept it. On the
other hand, I wanted to make some different plots from the ones in the example
analysis, including some that required loading prepared data. To check out
everything I did, see
[here](https://github.com/teddygroves/bibat/blob/main/bibat/examples/baseball/baseball/investigate.ipynb).

# Choosing priors using push-forward calibration

The trickiest thing about my analysis was setting prior distributions for the
parameters $k$ and $\sigma$ in the generalised Pareto models. To choose some
more or less plausible values I did a few prior-mode model runs and checked the
distributions of the resulting `alpha` variables. I wanted to make sure that
they all lay in the range corresponding to batting averages between about 0.1
and a little over 0.4. Here is a graph that shows the 1% to 99% prior quantiles
for each player's latent success percentage in both datasets alongside their
actually realised success rates.

![](../plots/prior_quantiles.png)

# Extending the analysis to the baseballdatabank data

To model the more recent data, all I had to do was create some new inference
folders with appropriate `prepared_data_dir` fields in their `config.toml`
files. For example, here is the `config.toml` file for the `gparetobdb`
inference:

```{.toml include="../inferences/gparetobdb/config.toml"}
```

After running `make analysis` one more time, I went back to the notebook
`baseball/investigate.ipynb` and made plots of both models' posterior 1%-99%
success rate intervals for both datasets:

![](../plots/posterior_quantiles.png)

I think this is very interesting. Both models' prior distributions had similar
regularisation levels, and they more or less agree about the abilities of the
players with the most at-bats, both in terms of locations and the widths of the
plausible intervals for the true success rate. However, the models ended up with
dramatically different certainty levels about the abilities of players with few
at-bats. This pattern was true both for the small 2006 dataset and the much
larger baseballdatabank dataset.

# Documenting the analysis

The final step was to document my analysis. To do this I edited the file
`docs/report.qmd`, then ran `quarto render docs/report.qmd`, which produced the
very HTML document that you are probably reading now! You can find the complete
`report.qmd` file
[here](https://github.com/teddygroves/bibat/blob/main/bibat/examples/baseball/docs/report.qmd).