### 2023-12-15 paper/jmlr_cover_letter.qmd
start writing jlmr paper

---
title: Cover letter - Bibat, Batteries-included Bayesian analysis template
---

Dear editors

This paper is intended for the machine learning open source software section
of JLMR.

All the authors are aware of this submission and consent to its review by JMLR.

Please find below all the information listed on the relevant JMLR page here
<https://www.jmlr.org/mloss/mloss-info.html>.

Best wishes,
Teddy Groves

## Open source licence

MIT licence: <https://github.com/teddygroves/bibat/LICENSE>

## Web address of the project

<https://github.com/teddygroves/bibat/>

## Software version to be reviewed

XXX VERSION XXX

## Evidence of active user community

Bibat has XXX GitHub stars and XXX active devlopers. In addition, bibat is a
PyOpenSci affiliated package: the PyOpenSci community provides rigorous peer
review (though not publication), publicity, community engagement and help with
software maintenance.

## Suggested reviewers

The following people have already reviewed bibat for PyOpenSci and therefore
might find it easy to provide another review:

- Meenal Jhajharia <meenal@mjhajharia.com>
- Oriol Abril Plaza <https://github.com/OriolAbril>
- Alexey Izmailov <alizma@brown.edu>

Editor: @mjhajharia
Reviewer 1: @OriolAbril
Reviewer 2: @alizma