### 2023-12-15 paper/main.qmd

---
title: "Bibat: Batteries-include Bayesian Analysis Template"
keywords: Bayesian inference,
abstract: >
  Despite their suitability for many scientific problems and the existence of
  sound theoretical and computational frameworks, adoption of Bayesian workflow
  in science remains a challenge. One reason for the difficulty is that it is
  difficult to write software that implements non-trivial Bayesian workflows.
  We aimed to address this difficulty by developing Bibat, a Python package
  providing  an interactive template that targets Bayesian statistical analysis
  projects. Bibat trivialises the otherwise painful process of structuring such
  a project and integrating together tools that help with individual parts of a
  Bayesian workflow, without compromising flexibility, scalability,
  interoperability or reproducibility. Bibat is available on the Python
  Package index,  documented at <https://bibat.readthedocs.io/> and
  developed at <https://github.com/teddygroves/bibat/>.
number-sections: true
author:
  - name: "Teddy Groves"
    orcid: 0000-0002-7109-3270
    email: tedgro@biosustain.dtu.dk
    affiliations:
      - id: biosustain
        name: DTU
        department: The Novo Nordisk Foundation Center for Biosustainability
        city: Kongens Lyngby
        country: Denmark
    attributes:
      corresponding: true
bibliography: bibliography.bib
format:
  html:
    embed-resources: true
  pdf:
    keep-tex: true
---

Many areas of science would benefit from replacing traditional statistical
methodologies with Bayesian workflow, as described in, for example
[@gelmanBayesianWorkflow2020;@grinsztajnBayesianWorkflowDisease2021;
@gabryVisualizationBayesianWorkflow2019].

There are now software tools that address most individual aspects of a Bayesian
workflow, from fetching and manipulating data to specifying, computing, storing,
inspecting and documenting statistical inferences. However, writing a software
project implementing a Bayesian workflow remains challenging due partly to the
difficulty of orchestrating and configuring these elements.

Bibat is a Python package that addresses this difficulty by providing an
interactive template for Bayesian workflow projects.

# Installation and usage

Bibat is installed by running the command `pip install bibat` and then used by
running the command `bibat`.

This command triggers an interactive form which prompts the user for
configuration information including project and repository name, author name, a
short description and choices of open source licence options, documentation
formats and whether to include tests and continuous integration.

Bibat then creates a folder with the chosen repository name in the current
working directory, containing code that implements bibat's example analysis.

# Documentation

Bibat is documented at <https://github.com/teddygroves/bibat/>. The
documentation website includes instructions for getting started, a detailed
explanation of bibat's concepts, an extended vignette illustrating intended
usage, full description of the python API and command line interface,
instructions for contributing and a section discussing accessibility
considerations.

Bibat's documentation website is continuously checked to ensure that it
implements accessibility best practices.

# Software

Bibat is implemented using Python 3. Its main dependencies are cookiecutter
[@greenfeldCookiecutter2021], pydantic [@pydanticdevelopersPydantic2022] and
click [@clickdevelopersClickPythonComposable2022].

Bibat projects have the following dependencies:

* Python 3 [@vanrossumPythonReferenceManual2009] and standard scientific Python
packages for data manipulation
* Stan [@carpenterStanProbabilisticProgramming2017] for specifying statistical
models and performing inferences
* cmdstanpy [@standevelopmentteamCmdStanPy2022] for interfacing between Python
and Stan
* arviz [@kumarArviZUnifiedLibrary2019] for storing and analysing completed
inferences
* pydantic and pandera [@niels_bantilan-proc-scipy-2020] for validation
* make [@stallman1991gnu] for automation.
 * Sphinx [@georgbrandlandthesphinxteamSphinx2022] and Quarto
[@Allaire_Quarto_2022] for documentation.

Bibat is continuously tested to ensure that it works on the operating systems
Linux, macOS and Windows.

Bibat's continuous integration runs on windows and ubuntu and runs bibat's test
suite as well as an end-to-end functional test on all supported Python versions.

# How bibat addresses specific Bayesian workflow difficulties

This section describes some specific problems that often affect Bayesian
workflow projects and which bibat helps to ameliorate.

## Data modelling

Any statistical analysis requires a definition of a prepared dataset, whether
implicitly or explicitly. Ideally the definition is explicit and allows for
arbitrarily many prepared datasets, accommodates both tabular and non-tabular
formats and provides functionality like validation and serialisation.

Bibat achieves this goal by providing abstract models for prepared data,
preconfigured to work well together, integrated into a data preparation pipeline
and able to be easily customised to suit any analysis.

Without a template such as bibat, users must either devise a new data modelling
framework from scratch as part of an analysis, which is time-consuming, or
else to miss out on either the flexibility or features that bibat's approach
provides.

## Reproducibility

Bibat ensures reproducibility by providing a preconfigured makefile with a
target `analysis` triggering creation of an isolated environment, installation
of dependencies, data preparation, statistical computation and analysis of
results. In this way a bibat analysis can be reproduced on most platforms
using a single command.

Bibat also provides its Python code in the form of a package configured using
modern conventions for specifying dependencies and configuring tooling, so that
it is easy to maintain reproducibility as the analysis develops.

## Collaboration

Bibat provides a preconfigured test environment, continuous integration, linting
and pre-commit hooks, making it suitable for collaborative software development.
In addition, including documentation as a first class component of the analysis
addresses a common problem in academic statistics projects where the paper gets
out of sync with the code.

## Installing dependencies

Bibat's makefile detects the current operating system and attempts to install
cmdstan appropriately if necessary. This functionality addresses a common
issue where researchers find it difficult to install Stan, especially on Windows.

## Fitting modes

As part of a Bayesian workflow it is often necessary to fit a model and dataset
in different ways. For example, one might perform MCMC sampling of both the
prior and posterior distributions, perform multiple leave-out-one-fold fits for
cross-validation or need to compare MCMC sampling with an optimisation-based
alternative.

Bibat accommodates this ubiquitous scenario by introducing an abstraction called
"fitting mode" and a corresponding Pydantic base class `FittingMode`, along with
and several subclasses for commonly used cases.

This abstraction allows bibat projects to handle fitting a model and dataset in
different ways appropriately and flexibly. For example, the provided prior
sampling fitting mode creates a Stan input dictionary with the `likelihood`
data variable set to `0`, performs MCMC sampling and writes data to the
`InferenceData` group `prior`.

## Configure fitting concisely.

Bibat provides a Pydantic class `InferenceConfiguration`, parsed from toml files
in the `inferences` directory. By writing a toml file per inference the user can
configure statistical model, dataset, fitting modes and computational settings
(globally or per fitting mode).

This approach appropriately separates configuration from logic and avoids
unnecessary duplication of configuration for runs with the same model and
dataset.

## Storing MCMC samples

Bibat provides code that automatically saves its output `InferenceData` objects
in zarr format. This format distributes information from the same set of samples
into multiple smaller files, thereby helping to avoid file size limits of online
repository hosting services.

# Comparison with alternative software

Other than bibat, there is currently no interactive template that specifically
targets Bayesian workflow projects. There are some templates that arguably
encompass Bayesian workflow as a special case of data analysis project, such
as cookiecutter-data-science [@drivendataCookiecutterdatascience2022], but
these are of limited use compared with a specialised template due to the many
specificities of Bayesian workflow.

There is some software that addresses the general task of facilitating Bayesian
workflow, but differently from bibat. For example, bambi [@capretto2020]
and brms [@burknerBrmsPackageBayesian2017] aim to make implementing Bayesian
workflows easier by trivialising the task of specifying a Bayesian generalised
linear model and fitting it to a single tabular dataset. This approach does not
address several difficulties that bibat does address,

# References