### 2024-01-29 paper/main.qmd
Update paper

---
title: "Bibat: Batteries-include Bayesian Analysis Template"
abstract: >
  Bayesian statistical workflow offers a powerful way to learn from data,
  but software software projects that implement complex Bayesian workflows
  in practice are unusual, partly due to the difficulty of orchestrating
  Bayesian statistical software. Bibat addresses this challenge by providing a
  full-featured, scalable Bayesian statistical analysis project
  using an interactive template. Bibat is available on the Python Package index,
  documented at <https://bibat.readthedocs.io/> and developed at
  <https://github.com/teddygroves/bibat/>. This paper explains the motivation
  for bibat, briefly describes intended usage, discusses key design choices,
  and reviews several examples of bibat's use in scientific applications.
number-sections: true
author:
  - name: "Teddy Groves"
    orcid: 0000-0002-7109-3270
    email: tedgro@biosustain.dtu.dk
    affiliations:
      - id: 1
        name: Danish Technical University
        department: The Novo Nordisk Foundation Center for Biosustainability
        city: Kongens Lyngby
        country: Denmark
    attributes:
      corresponding: true

bibliography: bibliography.bib

format:
  html:
    embed-resources: true
  acm-pdf:
    keep-tex: true

acm-metadata:
  # comment this out to make submission anonymous
  # anonymous: true

  # comment this out to build a draft version
  final: true

  # comment this out to specify detailed document options
  acmart-options: sigconf, review

  # acm preamble information
  copyright-year: 2024
  acm-year: 2024
  copyright: acmcopyright
  # doi: XXXXXXX.XXXXXXX
  conference-acronym: "KDD"
  conference-name: "KDD"
  conference-date: Aug 25--29, 2024
  conference-location: Barcelona, Esp
  keywords:
    - Python
    - Bayesian workflow
    - Methodology
  teaser:
    image: img/workflow.png
    caption: |
      Schematic representation of a Bayesian workflow implemented using bibat.
      The author inspects their analysis's results, edits files corresponding
      to the boxes on the left, runs the command `make analysis`, then repeats.
      Note that this workflow is modular, accommodates plurality and results in
      a final analysis that can be fully reproduced using a single command.
    description: "DESCRIPTION"

---

# Introduction: the problem of orchestrating Bayesian workflow software

The idea that Bayesian statistical analysis comprises not just
inference, but also specific approaches to related activities like data
preparation, model design, diagnosis, debugging and criticism, dates
at least to @boxBayesianInferenceStatistical1992. There is increasing
scholarly recognition of the need for a holistic view of "Bayesian
workflow" [@gelmanBayesianWorkflow2020;@grinsztajnBayesianWorkflowDisease2021;
@gabryVisualizationBayesianWorkflow2019]. Relatedly, software tools now
exist that address most individual aspects of a Bayesian workflow: see
@strumbeljPresentFutureSoftware for a review of the state of the art.

Unfortunately, currently available tools typically address one, or at most
a handful, of Bayesian workflow activities; it is left to the individual
project team to orchestrate the components. Writing software that performs
this orchestration can be time-consuming and tricky, especially in the common
scenario where it is not initially clear how many, or what kind of, statistical
models, datasets, data manipulations or investigations an analysis will require.

Bibat addresses this difficulty by providing a full-featured, high-quality
Bayesian workflow project that can be extended to implement a wide range of
statistical analyses.

# How bibat works

## Installation and usage

Bibat can be installed on the Windows, linux or macos command line on by running
the command `pip install bibat` and then used by running the command `bibat`.
This command triggers an interactive form which prompts the user to select a
range of customisation options. Bibat then creates a new directory containing
code that implements an example analysis, with customisations reflecting the
user's choices. This analysis works immediately, and can be reproduced with
the single command `make analysis` without the need for any further action by
the user: in this sense bibat comes with "batteries included".

## Documentation

Bibat is documented at <https://github.com/teddygroves/bibat/>. The
documentation website includes instructions for getting started, a detailed
explanation of bibat's concepts and an extended vignette illustrating how to
implement a complex statistical analysis starting from bibat's example analysis
usage. In addition, the documentation site contains a full description of
bibat's python API and command line interface, instructions for contributing and
a section discussing accessibility considerations.

# Design choices

Bibat's design was informed by these considerations.

## Accommodate a wide range of statistical analyses

Bibat projects explicitly allow for plurality at the level of input datasets,
prepared datasets, statistical models, fitting modes, computation methods and
analyses. In addition, bibat ensures that there are minimal restrictions on the
kind of components: for example, datasets need not be singular or tabular, and
statistical models need not be representable in formula syntax.

Thanks to these accommodations a project team using bibat should typically not
need to foresee the ultimate requirements of their analysis before starting
the project.

## Encourage reproducibility

Bibat provides a preconfigured makefile with a target `analysis` triggering
creation of an isolated environment, installation of dependencies, data
preparation, statistical computation and analysis of results. In this way a
bibat analysis can be reproduced on most platforms using a single command.

Bibat also provides its Python code in the form of a package configured using
modern conventions for specifying dependencies and configuring tooling, so that
it is easy to maintain reproducibility as the analysis develops.

## Use widely-adopted, open source and actively developed tools

Bibat projects are is written in modern Python and uses pydantic
@pydanticdevelopersPydantic2022 and pandera [@niels_bantilan-proc-scipy-2020]
for data modelling, Stan [@carpenterStanProbabilisticProgramming2017] for
statistical inference, cmdstanpy [@standevelopmentteamCmdStanPy2022] for
python-Stan interface, arviz [@kumarArviZUnifiedLibrary2019] for storing and
analyzing inferences and sphinx [@georgbrandlandthesphinxteamSphinx2022] and
quarto [@Allaire_Quarto_2022] for documentation.

Bibat itself is also written in modern Python and uses the popular
tools cookiecutter [@greenfeldCookiecutter2021], pydantic and click
[@clickdevelopersClickPythonComposable2022].

Bibat's makefile detects the current operating system and attempts to install
cmdstan appropriately if necessary. This functionality addresses a common
issue where researchers find it difficult to install Stan, especially on Windows.

## Implement community standards and best practices for collaborative software development

Bibat projects include a preconfigured test environment, continuous integration,
linting and pre-commit hooks, making them suitable for collaborative software
development. In addition, including documentation as a first class component of
the analysis addresses a common problem in academic statistics projects where
the paper gets out of sync with the code.

Bibat itself is continuously tested to ensure that it works on the operating
systems Linux, macOS and Windows. Bibat's continuous integration runs a test
suite as well as an end-to-end functional test on all supported Python versions.

Bibat is part of the PyOpenSci ecosystem, allowing for community help with
maintenance as well as peer review for code and documentation quality, usability
and accessibility. The PyOpenSci peer review for bibat can be found here: <https://github.com/pyOpenSci/software-submission/issues/83>

# Challenges

This section describes some specific challenges that often affect Bayesian
workflow projects and how bibat addresses them.

## Address complexity using a modular, file based approach

As discussed in @gelmanBayesianWorkflow2020, Bayesian workflows are complicated,
featuring plurality, cyclicity and complexity at many levels. Bibat accommodates
this by separating non-interacting analysis components into modules and by
serialising data to files wherever possible. Prepared datasets, statistical
models, inference configurations, inference results, plots and analyses all
have file representations. Fitting modes, data manipulations and data models are
modularised in code through the use of appropriately structured data classes
and functions.

Thanks to this approach it is possible to perform small sub analyses
individually and to iteratively add components without needing to consider
everything at once.

## Fitting modes

As part of a Bayesian workflow it can be necessary to fit a model and dataset
in different ways. For example, one might perform MCMC sampling of both the
prior and posterior distributions, perform multiple leave-out-one-fold fits for
cross-validation or need to compare MCMC sampling with an optimisation-based
alternative.

Bibat accommodates this need by introducing an abstraction called "fitting
mode". This abstraction allows bibat projects to handle fitting a model
and dataset in different ways appropriately and flexibly. For example, the
provided prior sampling fitting mode creates a Stan input dictionary with the
`likelihood` data variable set to `0`, performs MCMC sampling and writes data
to the `InferenceData` group `prior`. Bibat provides fitting modes corresponding
for prior sampling, posterior sampling and k-fold posterior sampling. Users can
easily add additional fitting modes by copying these examples.

# Post launch use

Bibat has been used for several real statistical analyses.

@grovesteddyDgfreg2023 used bibat to compare a Bayesian and two non-Bayesian
approaches to modelling a biochemical thermodynamics dataset. Bibat facilitated
this analysis even though it was not very large by providing the fitting mode
abstraction, which was very useful for comparing the different methods.

In @grovesBaseball2022, Bibat was used to implement a sports analysis involving
two datasets, two models and four inferences, demonstrating that the generalised
Pareto distribution can be used to describe hitting ability in baseball. This
analysis is now included in bibat as an illustration, along with an accompanying
tutorial.

:::{#fig-baseball}

![](img/baseball.png){width=50%}

A graphical posterior predictive check produced as part of a bibat
analysis that fit two statistical models to two datasets of baseball data.
The coloured lines show each model's posterior predictive distributions
and the black dots show the two observed datasets.  See <https://github.com/teddygroves/bibat/tree/main/bibat/examples/baseball>
for the full analysis.

:::

In @grovesteddySphincter2024, bibat was used to implement an analysis of
cerebrovascular data from mice, involving two raw datasets, 6 prepared
datasets, 15 models and 15 inferences. In this case bibat's modular design made
it straightforward to iteratively add data transformations and models while
maintaining reproducibility.

These cases illustrate that bibat can be useful in a variety of real Bayesian
workflows, with different sizes, subject matters and emphases.

Bibat has a growing user community, with 16 GitHub stars at the time of writing.

# Comparison with alternative software

Other than bibat, there is currently no interactive template that specifically
targets Bayesian workflow projects. There are some templates that arguably
encompass Bayesian workflow as a special case of data analysis project, such
as cookiecutter-data-science [@drivendataCookiecutterdatascience2022], but
these are of limited use compared with a specialised template due to the many
specificities of Bayesian workflow.

There is some software that addresses the general task of facilitating
Bayesian workflow, but using a different approach from bibat's. For example,
bambi [@capretto2020] and brms [@burknerBrmsPackageBayesian2017] aim to make
implementing Bayesian workflows easier by providing ergonomic ways to specify
and fit Bayesian regression models to tabular datasets. Bibat is complementary
with these packages, as it targets use cases that they do not support, such as
analyses where complex datasets or custom models might be required.

# Statements and Declarations

The authors have no financial or non-financial interests that are directly or
indirectly related to the work submitted for publication.

# References