### 2024-01-30 paper/main.qmd
Update paper

---
title: "Bibat: Batteries-include Bayesian Analysis Template"
abstract: >
  Bayesian statistical workflow offers a powerful way to learn from data,
  but software software projects that implement complex Bayesian workflows
  in practice are unusual, partly due to the difficulty of orchestrating
  Bayesian statistical software. Bibat addresses this challenge by providing
  a full-featured, scalable Bayesian statistical analysis project using
  an interactive template. Bibat is available on the Python Package index,
  documented at <https://bibat.readthedocs.io/> and developed at <https://github.com/teddygroves/bibat/>. This paper explains the motivation for bibat,
  describes intended usage, discusses bibat's design, compares bibat with similar software, highlights several examples of bibat's use in science and provides links to community resources associated with bibat.
number-sections: true
author:
  - name: "Teddy Groves"
    orcid: 0000-0002-7109-3270
    email: tedgro@biosustain.dtu.dk
    affiliations:
      - id: 1
        name: Danish Technical University
        department: The Novo Nordisk Foundation Center for Biosustainability
        city: Kongens Lyngby
        country: Denmark
    attributes:
      corresponding: true
bibliography: bibliography.bib
keywords: [Bayesian workflow, Methodology, Software]
format:
  typst: default
  html:
    embed-resources: true
  latex: default
---

# Introduction: the problem of orchestrating Bayesian workflow software

The idea that Bayesian statistical analysis comprises not just
inference, but also specific approaches to related activities like data
preparation, model design, diagnosis, debugging and criticism, dates
at least to @boxBayesianInferenceStatistical1992. There is increasing
scholarly recognition of the need for a holistic view of "Bayesian
workflow" [@gelmanBayesianWorkflow2020;@grinsztajnBayesianWorkflowDisease2021;
@gabryVisualizationBayesianWorkflow2019]. Relatedly, software tools now
exist that address most individual aspects of a Bayesian workflow: see
@strumbeljPresentFutureSoftware for a review of the state of the art.

Unfortunately, currently available tools typically address one, or at most
a handful, of Bayesian workflow activities; it is left to the individual
project team to orchestrate the components. Writing software that performs
this orchestration can be time-consuming and tricky, especially in the common
scenario where it is not initially clear how many, or what kind of, statistical
models, datasets, data manipulations or investigations an analysis will require.

Bibat addresses this difficulty by providing a full-featured, high-quality
Bayesian workflow project that can be extended to implement a wide range of
statistical analyses.

# How bibat works

## Installation and usage

Bibat can be installed on the Windows, Linux or macOS command line on by running
the command `pip install bibat` and then used by running the command `bibat`.
This command triggers an interactive form which prompts the user to select a
range of customisation options. Bibat then creates a new directory containing
code that implements an example analysis, with customisations reflecting the
user's choices. This analysis works immediately, and can be reproduced with
the single command `make analysis` without the need for any further action by
the user: in this sense bibat comes with batteries included.

@fig-workflow illustrates the components of a bibat-based Bayesian workflow and
shows how it proceeds: the project team edits the code components, then runs
`make analysis` triggering creation of the result components. After inspecting
these they repeat the process, leading to a cycle that ultimately results in a
complete, easily reproducible analysis.

:::{#fig-workflow}

![](img/workflow.png){width=100%}

Schematic representation of a Bayesian workflow implemented using bibat. The
author inspects their analysis's results, edits code corresponding to the
boxes on the left, runs the command `make analysis`, then repeats. The diagram
illustrates several key features of bibat: inference components are modular and
plural, the overall workflow is iterative and cyclical and the whole analysis
can be executed with a single command.

:::

## Documentation

Bibat is documented at <https://github.com/teddygroves/bibat/>. The
documentation website includes instructions for getting started, a detailed
explanation of bibat's concepts and an extended vignette illustrating how to
implement a complex statistical analysis starting from bibat's example analysis
usage. In addition, the documentation site contains a full description of
bibat's python API and command line interface, instructions for contributing and
a section discussing accessibility considerations.

# Design choices

Bibat's design was informed by the aims to accommodate the many sources of
complexity and in a Bayesian workflow project, to ensure easy reproducibility
and to integrate many open-source, widely-adopted and powerful Bayesian workflow
tools and to encourage collaborative development.

As discussed in @gelmanBayesianWorkflow2020, Bayesian workflows are complicated,
featuring plurality, cyclicity and complexity at many levels. As a specialised
Bayesian workflow template, a key goal for bibat was to manage this complexity.
Bibat achieves this aim by separating non-interacting analysis components into
separate, potentially plural modules and by serialising data to files wherever
possible. Prepared datasets, statistical models, inference configurations,
inference results, plots and analyses all have file representations. Fitting
modes, data manipulations and data models are modularised in code through the
use of appropriately structured data classes and functions. Thanks to this
modular approach it is possible to perform small sub analyses individually
and to iteratively expand the analysis by adding components without needing to
consider everything at once. In addition, bibat ensures that there are minimal
restrictions on the components: for example, datasets need not be singular or
tabular and statistical models need not be representable in formula syntax.
Thanks to these accommodations a project team using bibat should typically not
need to foresee the ultimate requirements of their analysis before starting
the project.

Bibat encourages reproducibility mainly by providing a preconfigured makefile
with a target `analysis` that triggers creation of an isolated environment,
installation of dependencies, data preparation, statistical computation and
analysis of results. In this way a bibat analysis can be reproduced on most
platforms using a single command. In particular, this target attempts to install
cmdstan if necessary, using a recipe tailored to the host operating system.
This functionality addresses a common issue where researchers find it difficult
to install Stan, especially on Windows. A second way in which bibat encourages
reproducibility is by providing its Python code in the form of a preconfigured
package. Thanks to the use of modern conventions for specifying dependencies
and configuring tooling, we expect that this will make bibat analyses easier
for other researchers to work with, thereby facilitating replication while also
making initial development more convenient.

Bibat integrates many widely-adopted, open-source and powerful tools to
implement the components of a Bayesian workflow. Bibat projects are is
written in modern Python and uses pydantic @pydanticdevelopersPydantic2022
and pandera [@niels_bantilan-proc-scipy-2020] for data modelling, Stan
[@carpenterStanProbabilisticProgramming2017] for statistical inference,
cmdstanpy [@standevelopmentteamCmdStanPy2022] for python-Stan interface, arviz
[@kumarArviZUnifiedLibrary2019] for storing and analyzing inferences and sphinx
[@georgbrandlandthesphinxteamSphinx2022] and quarto [@Allaire_Quarto_2022]
for documentation.  Bibat itself is also written in modern Python and uses the
popular tools cookiecutter [@greenfeldCookiecutter2021], pydantic and click
[@clickdevelopersClickPythonComposable2022].

To encourage collaborative development of Bayesian workflow projects, Bibat
projects include a preconfigured test environment, continuous integration,
linting and pre-commit hooks. In addition, including documentation as a
first class component of the analysis addresses a common problem in academic
statistics projects where the paper gets out of sync with the code. Bibat itself
is continuously tested to ensure that it works on the operating systems Linux,
macOS and Windows. Bibat's continuous integration runs a test suite as well as
an end-to-end functional test on all supported Python versions.

# Fitting modes

The most novel part of bibat's design is the introduction of an abstraction
called "fitting mode", which allows bibat project to handle different ways
of fitting a model to a dataset appropriately and flexibly.  This is often
necessary as part of a Bayesian workflow: for example, one might perform
MCMC sampling of both the prior and posterior distributions, perform multiple
leave-out-one-fold fits for cross-validation or need to compare MCMC sampling
with an optimisation-based alternative.

Fitting modes in bibat projects take the form of instances of the class
`FittingMode`. Each fitting mode contains a name, a function that fits a
prepared dataset and instructions for how and where to save the results.
For example, the provided prior sampling fitting mode creates a Stan input
dictionary with the `likelihood` data variable set to `0`, performs MCMC
sampling and writes data to the `InferenceData` group `prior`. This design
allows multiple fits to be stored alongside each other naturally, facilitating
many common Bayesian workflow analyses, and can easily be configured to prevent
unnecessary computations.

Bibat provides fitting modes corresponding for prior sampling, posterior
sampling and k-fold posterior sampling. Users can easily add additional fitting
modes by copying these examples, or modify the `FittingMode` class to achieve
even richer functionality.

# Comparison with alternative software

Other than bibat, there is currently no interactive template that specifically
targets Bayesian workflow projects. There are some templates that arguably
encompass Bayesian workflow as a special case of data analysis project, such
as cookiecutter-data-science [@drivendataCookiecutterdatascience2022], but
these are of limited use compared with a specialised template due to the many
specificities of Bayesian workflow.

There is some software that addresses the general task of facilitating
Bayesian workflow, but using a different approach from bibat's. For example,
bambi [@capretto2020] and brms [@burknerBrmsPackageBayesian2017] aim to make
implementing Bayesian workflows easier by providing ergonomic ways to specify
and fit Bayesian regression models to tabular datasets. Bibat is complementary
with these packages, as it targets use cases that they do not support, such as
analyses where complex datasets or custom models might be required.

# Case studies

The following cases illustrate how bibat has been used in practice to facilitate
Bayesian workflow projects.

@grovesteddyDgfreg2023 used bibat to compare a Bayesian and two non-Bayesian
approaches to modelling a biochemical thermodynamics dataset. Bibat facilitated
this analysis even though it was not very large---the final analysis contained
one dataset, three models and three inferences---because the fitting mode
abstraction allowed for straightforward comparison of the different methods.
Additionally, bibat made it easier to iteratively investigate and discard models
that did not form part of the final analysis.

In @grovesBaseball2022, Bibat was used to implement a sports analysis involving
two datasets, two models and four inferences, demonstrating that the generalised
Pareto distribution can be used to describe hitting ability in baseball. This
analysis is now included in bibat as an illustration, along with an accompanying
tutorial. An illustrative graphic from this analysis is shown in @fig-baseball.

In this case bibat was useful because of its ability to implement arbitrary
statistical models, as latent generalised Pareto distributions are not supported
by generic regression packages like brms and bambi. Further, bibat's modular
design made it easier to implement this medium-sized analysis with two datasets,
two models and six inferences.

:::{#fig-baseball}

![](img/baseball.png){width=100%}

A graphical posterior predictive check produced as part of a bibat
analysis that fit two statistical models to two datasets of baseball data.
The coloured lines show each model's posterior predictive distributions
and the black dots show the two observed datasets.  See <https://github.com/teddygroves/bibat/tree/main/bibat/examples/baseball>
for the full analysis.

:::

In @grovesteddySphincter2024, bibat was used to implement a large analysis
of cerebrovascular data from mice, involving two raw datasets, 6 prepared
datasets, 15 models and 15 inferences. Again it was not possible to implement
the analysis using formula-based packages due to the need for custom modelling
approaches to take into account expert knowledge about both the target system
and the measurement apparatus. Due to the number of datasets and models that
needed to be considered, implementing this analysis without using bibat or a
similar structured approach would have required a combination of more time and/
or compromises in software quality.

These cases illustrate that bibat can be useful in a variety of real Bayesian
workflows, with different sizes, subject matters and emphases.

# Community

Bibat is developed in public and encourages community contribution. Please see the contributing page <https://github.com/teddygroves/bibat/blob/main/CONTRIBUTING.rst> and code of conduct <https://github.com/teddygroves/bibat/blob/main/CODE_OF_CONDUCT.md> if you would like to help develop bibat.

Bibat has a growing user community, with 16 GitHub stars at the time of
writing, and is affiliated with cmdstanpy through a link on its documentation
website. Bibat is also affiliated with the Python scientific software community
PyOpenSci, allowing for help with maintenance as well as peer review for
code and documentation quality, usability and accessibility. The PyOpenSci
peer review for bibat can be found here: <https://github.com/pyOpenSci/software-submission/issues/83>.

# Author contributions

# Acknowledgements