### 2024-02-02 paper/main.tex
copier

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  letterpaper,
  DIV=11,
  numbers=noendperiod]{scrartcl}

\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else
    % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{5}
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
% definitions for citeproc citations
\NewDocumentCommand\citeproctext{}{}
\NewDocumentCommand\citeproc{mm}{%
  \begingroup\def\citeproctext{#2}\cite{#1}\endgroup}
\makeatletter
 % allow citations to break across lines
 \let\@cite@ofmt\@firstofone
 % avoid brackets around text for \cite:
 \def\@biblabel#1{}
 \def\@cite#1#2{{#1\if@tempswa , #2\fi}}
\makeatother
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newenvironment{CSLReferences}[2] % #1 hanging-indent, #2 entry-spacing
 {\begin{list}{}{%
  \setlength{\itemindent}{0pt}
  \setlength{\leftmargin}{0pt}
  \setlength{\parsep}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
   \setlength{\leftmargin}{\cslhangindent}
   \setlength{\itemindent}{-1\cslhangindent}
  \fi
  % set entry spacing
  \setlength{\itemsep}{#2\baselineskip}}}
 {\end{list}}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{\hfill\break\parbox[t]{\linewidth}{\strut\ignorespaces#1\strut}}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

\KOMAoption{captions}{tableheading}
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage{bookmark}

\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Bibat: Batteries-include Bayesian Analysis Template},
  pdfauthor={Teddy Groves},
  pdfkeywords={Bayesian workflow, Methodology, Software},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\title{Bibat: Batteries-include Bayesian Analysis Template}
\author{Teddy Groves}
\date{}

\begin{document}
\maketitle
\begin{abstract}
Bayesian statistical workflow offers a powerful way to learn from data,
but software software projects that implement complex Bayesian workflows
in practice are unusual, partly due to the difficulty of orchestrating
Bayesian statistical software. Bibat addresses this challenge by
providing a full-featured, scalable Bayesian statistical analysis
project using an interactive template. Bibat is available on the Python
Package index, documented at \url{https://bibat.readthedocs.io/} and
developed at \url{https://github.com/teddygroves/bibat/}. This paper
explains the motivation for bibat, describes intended usage, discusses
bibat's design, compares bibat with similar software, highlights several
examples of bibat's use in science and provides links to community
resources associated with bibat.
\end{abstract}

\section{Introduction: the problem of orchestrating Bayesian workflow
software}\label{introduction-the-problem-of-orchestrating-bayesian-workflow-software}

The term ``Bayesian workflow'' captures the idea that Bayesian
statistical analysis comprises not just inference, but also specific
approaches to related activities like data preparation, model design,
diagnosis, debugging and criticism. This idea can be found in Box and
Tiao (1992) and has recently received increasing scholarly recognition
of (Gelman et al. 2020; Grinsztajn et al. 2021; Gabry et al. 2019).
Software tools now exist that address most individual aspects of a
Bayesian workflow: see Štrumbelj et al. (2024) for a review of the state
of the art.

Unfortunately, each tool typically addresses one, or at most a few, of
the many activities that comprise a real Bayesian workflow software
project; it is left to the individual project team to orchestrate all of
the tools they require. Writing software that performs this
orchestration can be time-consuming and tricky, especially in the common
scenario where it is not initially clear how many, or what kind of,
statistical models, datasets, data manipulations or investigations an
analysis will require.

Bibat is a new tool that addresses the difficulty of orchestrating
Bayesian workflow software by providing a full-featured, high-quality
project that can be extended to implement a wide range of statistical
analyses.

\section{How bibat works}\label{how-bibat-works}

Bibat can be installed on the Windows, Linux or macOS command line on by
running the command \texttt{pip\ install\ bibat} and then used by
running the command \texttt{bibat}. This command triggers an interactive
form which prompts the user to select a range of customisation options.
Bibat then creates a new directory containing code that implements an
example analysis, with customisations reflecting the user's choices.
This analysis works immediately, and can be reproduced with the single
command \texttt{make\ analysis} without the need for any further action
by the user: in this sense bibat comes with batteries included.

Figure~\ref{fig-workflow} illustrates the components of a bibat-based
Bayesian workflow and shows how it proceeds: the project team edits the
code components, then runs \texttt{make\ analysis} triggering creation
of the result components. After inspecting these they repeat the
process, leading to a cycle that ultimately results in a complete,
easily reproducible analysis.

\begin{figure}

\centering{

\includegraphics[width=1\textwidth,height=\textheight]{img/workflow.png}

}

\caption{\label{fig-workflow}Schematic representation of a Bayesian
workflow implemented using bibat. The author inspects their analysis's
results, edits code corresponding to the boxes on the left, runs the
command \texttt{make\ analysis}, then repeats. The diagram illustrates
several key features of bibat: inference components are modular and
plural, the overall workflow is iterative and cyclical and the whole
analysis can be executed with a single command.}

\end{figure}%

Bibat is documented at \url{https://github.com/teddygroves/bibat/}. The
documentation website includes instructions for getting started, a
detailed explanation of bibat's concepts and an extended vignette
illustrating how to implement a complex statistical analysis starting
from bibat's example analysis usage. In addition, the documentation site
contains a full description of bibat's python API and command line
interface, instructions for contributing and a section discussing
accessibility considerations.

\section{Design choices}\label{design-choices}

Bibat's design was informed by the aims to accommodate the many sources
of complexity and in a Bayesian workflow project, to ensure easy
reproducibility and to integrate many open-source, widely-adopted and
powerful Bayesian workflow tools and to encourage collaborative
development.

As discussed in Gelman et al. (2020), Bayesian workflows are
complicated, featuring plurality, cyclicity and complexity at many
levels. As a specialised Bayesian workflow template, a key goal for
bibat was to manage this complexity. Bibat achieves this aim by
separating non-interacting analysis components into separate,
potentially plural modules and by serialising data to files wherever
possible. Prepared datasets, statistical models, inference
configurations, inference results, plots and analyses all have file
representations. Fitting modes, data manipulations and data models are
modularised in code through the use of appropriately structured data
classes and functions. Thanks to this modular approach it is possible to
perform small sub analyses individually and to iteratively expand the
analysis by adding components without needing to consider everything at
once. In addition, bibat ensures that there are minimal restrictions on
the components: for example, datasets need not be singular or tabular
and statistical models need not be representable in formula syntax.
Thanks to these accommodations a project team using bibat should
typically not need to foresee the ultimate requirements of their
analysis before starting the project.

Bibat encourages reproducibility mainly by providing a preconfigured
makefile with a target \texttt{analysis} that triggers creation of an
isolated environment, installation of dependencies, data preparation,
statistical computation and analysis of results. In this way a bibat
analysis can be reproduced on most platforms using a single command. In
particular, this target attempts to install cmdstan if necessary, using
a recipe tailored to the host operating system. This functionality
addresses a common issue where researchers find it difficult to install
Stan, especially on Windows. A second way in which bibat encourages
reproducibility is by providing its Python code in the form of a
preconfigured package. Thanks to the use of modern conventions for
specifying dependencies and configuring tooling, we expect that this
will make bibat analyses easier for other researchers to work with,
thereby facilitating replication while also making initial development
more convenient.

Bibat integrates many widely-adopted, open-source and powerful tools to
implement the components of a Bayesian workflow. Bibat projects are is
written in modern Python and uses pydantic Pydantic developers (2022)
and pandera (Niels Bantilan 2020) for data modelling, Stan (Carpenter et
al. 2017) for statistical inference, cmdstanpy (Stan Development Team
2022) for python-Stan interface, arviz (Kumar et al. 2019) for storing
and analyzing inferences and sphinx (Georg Brandl and the Sphinx team
2022) and quarto (Allaire et al. 2022) for documentation. Bibat itself
is also written in modern Python and uses the popular tools cookiecutter
(Greenfeld et al. 2021), pydantic and click (Click Developers 2022).

To encourage collaborative development of Bayesian workflow projects,
Bibat projects include a preconfigured test environment, continuous
integration, linting and pre-commit hooks. In addition, including
documentation as a first class component of the analysis addresses a
common problem in academic statistics projects where the paper gets out
of sync with the code. Bibat itself is continuously tested to ensure
that it works on the operating systems Linux, macOS and Windows. Bibat's
continuous integration runs a test suite as well as an end-to-end
functional test on all supported Python versions.

\section{Fitting modes}\label{fitting-modes}

The most novel part of bibat's design is the introduction of an
abstraction called ``fitting mode'', which allows bibat project to
handle different ways of fitting a model to a dataset appropriately and
flexibly. This is often necessary as part of a Bayesian workflow: for
example, one might perform MCMC sampling of both the prior and posterior
distributions, perform multiple leave-out-one-fold fits for
cross-validation or need to compare MCMC sampling with an
optimisation-based alternative.

Fitting modes in bibat projects take the form of instances of the class
\texttt{FittingMode}. Each fitting mode contains a name, a function that
fits a prepared dataset and instructions for how and where to save the
results. For example, the provided prior sampling fitting mode creates a
Stan input dictionary with the \texttt{likelihood} data variable set to
\texttt{0}, performs MCMC sampling and writes data to the
\texttt{InferenceData} group \texttt{prior}. This design allows multiple
fits to be stored alongside each other naturally, facilitating many
common Bayesian workflow analyses, and can easily be configured to
prevent unnecessary computations.

Bibat provides fitting modes corresponding for prior sampling, posterior
sampling and k-fold posterior sampling. Users can easily add additional
fitting modes by copying these examples, or modify the
\texttt{FittingMode} class to achieve even richer functionality.

\section{Comparison with alternative
software}\label{comparison-with-alternative-software}

Other than bibat, there is currently no interactive template that
specifically targets Bayesian workflow projects. There are some
templates that arguably encompass Bayesian workflow as a special case of
data analysis project, such as cookiecutter-data-science (Driven Data
2022), but these are of limited use compared with a specialised template
due to the many specificities of Bayesian workflow.

There is some software that addresses the general task of facilitating
Bayesian workflow, but using a different approach from bibat's. For
example, bambi (Capretto et al. 2020) and brms (Bürkner 2017) aim to
make implementing Bayesian workflows easier by providing ergonomic ways
to specify and fit Bayesian regression models to tabular datasets. Bibat
is complementary with these packages, as it targets use cases that they
do not support, such as analyses where complex datasets or custom models
might be required.

\section{Case studies}\label{case-studies}

The following cases illustrate how bibat has been used in practice to
facilitate Bayesian workflow projects.

Groves and Jooste (2023) used bibat to compare a Bayesian and two
non-Bayesian approaches to modelling a biochemical thermodynamics
dataset. Bibat facilitated this analysis even though it was not very
large---the final analysis contained one dataset, three models and three
inferences---because the fitting mode abstraction allowed for
straightforward comparison of the different methods. Additionally, bibat
made it easier to iteratively investigate and discard models that did
not form part of the final analysis.

In Groves (2022), Bibat was used to implement a sports analysis
involving two datasets, two models and four inferences, demonstrating
that the generalised Pareto distribution can be used to describe hitting
ability in baseball. This analysis is now included in bibat as an
illustration, along with an accompanying tutorial. An illustrative
graphic from this analysis is shown in Figure~\ref{fig-baseball}.

In this case bibat was useful because of its ability to implement
arbitrary statistical models, as latent generalised Pareto distributions
are not supported by generic regression packages like brms and bambi.
Further, bibat's modular design made it easier to implement this
medium-sized analysis with two datasets, two models and six inferences.

\begin{figure}

\centering{

\includegraphics[width=1\textwidth,height=\textheight]{img/baseball.png}

}

\caption{\label{fig-baseball}A graphical posterior predictive check
produced as part of a bibat analysis that fit two statistical models to
two datasets of baseball data. The coloured lines show each model's
posterior predictive distributions and the black dots show the two
observed datasets. See
\url{https://github.com/teddygroves/bibat/tree/main/bibat/examples/baseball}
for the full analysis.}

\end{figure}%

In Groves (2024), bibat was used to implement a large analysis of
cerebrovascular data from mice, involving two raw datasets, 6 prepared
datasets, 15 models and 15 inferences. Again it was not possible to
implement the analysis using formula-based packages due to the need for
custom modelling approaches to take into account expert knowledge about
both the target system and the measurement apparatus. Due to the number
of datasets and models that needed to be considered, implementing this
analysis without using bibat or a similar structured approach would have
required a combination of more time and/ or compromises in software
quality.

These cases illustrate that bibat can be useful in a variety of real
Bayesian workflows, with different sizes, subject matters and emphases.

\section{Community}\label{community}

Bibat is developed in public and encourages community contribution.
Please see the contributing page
\url{https://github.com/teddygroves/bibat/blob/main/CONTRIBUTING.rst}
and code of conduct
\url{https://github.com/teddygroves/bibat/blob/main/CODE_OF_CONDUCT.md}
if you would like to help develop bibat.

Bibat has a growing user community, with 16 GitHub stars at the time of
writing, and is affiliated with cmdstanpy through a link on its
documentation website. Bibat is also affiliated with the Python
scientific software community PyOpenSci, allowing for help with
maintenance as well as peer review for code and documentation quality,
usability and accessibility. The PyOpenSci peer review for bibat can be
found here:
\url{https://github.com/pyOpenSci/software-submission/issues/83}.

\section{Author contributions}\label{author-contributions}

\section*{Acknowledgements}\label{acknowledgements}
\addcontentsline{toc}{section}{Acknowledgements}

\phantomsection\label{refs}
\begin{CSLReferences}{1}{0}
\bibitem[\citeproctext]{ref-Allaire_Quarto_2022}
Allaire, J. J., Charles Teague, Carlos Scheidegger, Yihui Xie, and
Christophe Dervieux. 2022. {``Quarto.''}
\url{https://doi.org/10.5281/zenodo.5960048}.

\bibitem[\citeproctext]{ref-boxBayesianInferenceStatistical1992}
Box, George E. P., and George C. Tiao. 1992. \emph{Bayesian Inference in
Statistical Analysis}. Wiley classics library ed. A {Wiley-Interscience}
Publication. {New York}: {Wiley}.

\bibitem[\citeproctext]{ref-burknerBrmsPackageBayesian2017}
Bürkner, Paul-Christian. 2017. {``Brms: {An R} Package for {Bayesian}
Multilevel Models Using {Stan}.''} \emph{Journal of Statistical
Software} 80 (1): 1--28. \url{https://doi.org/10.18637/jss.v080.i01}.

\bibitem[\citeproctext]{ref-capretto2020}
Capretto, Tomás, Camen Piho, Ravin Kumar, Jacob Westfall, Tal Yarkoni,
and Osvaldo A. Martin. 2020. {``Bambi: {A} Simple Interface for Fitting
{Bayesian} Linear Models in {Python}.''}
\url{https://arxiv.org/abs/2012.10754}.

\bibitem[\citeproctext]{ref-carpenterStanProbabilisticProgramming2017}
Carpenter, Bob, Andrew Gelman, Matthew D. Hoffman, Daniel Lee, Ben
Goodrich, Michael Betancourt, Marcus Brubaker, Jiqiang Guo, Peter Li,
and Allen Riddell. 2017. {``Stan: {A Probabilistic Programming
Language}.''} \emph{Journal of Statistical Software} 76 (1): 1--32.
\url{https://doi.org/10.18637/jss.v076.i01}.

\bibitem[\citeproctext]{ref-clickdevelopersClickPythonComposable2022}
Click Developers. 2022. {``Click: {Python} Composable Command Line
Interface Toolkit.''} Pallets. \url{https://pypi.org/project/click/}.

\bibitem[\citeproctext]{ref-drivendataCookiecutterdatascience2022}
Driven Data. 2022. {``Cookiecutter-Data-Science.''}
\url{https://github.com/drivendata/cookiecutter-data-science/}.

\bibitem[\citeproctext]{ref-gabryVisualizationBayesianWorkflow2019}
Gabry, Jonah, Daniel Simpson, Aki Vehtari, Michael Betancourt, and
Andrew Gelman. 2019. {``Visualization in {Bayesian Workflow}.''}
\emph{Journal of the Royal Statistical Society Series A: Statistics in
Society} 182 (2): 389--402. \url{https://doi.org/10.1111/rssa.12378}.

\bibitem[\citeproctext]{ref-gelmanBayesianWorkflow2020}
Gelman, Andrew, Aki Vehtari, Daniel Simpson, Charles C. Margossian, Bob
Carpenter, Yuling Yao, Lauren Kennedy, Jonah Gabry, Paul-Christian
Bürkner, and Martin Modrák. 2020. {``Bayesian {Workflow}.''}
\emph{arXiv:2011.01808 {[}Stat{]}}, November.
\url{http://arxiv.org/abs/2011.01808}.

\bibitem[\citeproctext]{ref-georgbrandlandthesphinxteamSphinx2022}
Georg Brandl and the Sphinx team. 2022. {``Sphinx.''}
\url{https://www.sphinx-doc.org/}.

\bibitem[\citeproctext]{ref-greenfeldCookiecutter2021}
Greenfeld, Audrey Roy, Dainiel Roy Greenfeld, Raphael Pierzina, et al.
2021. {``Cookiecutter.''} \url{https://pypi.org/project/cookiecutter/}.

\bibitem[\citeproctext]{ref-grinsztajnBayesianWorkflowDisease2021}
Grinsztajn, Léo, Elizaveta Semenova, Charles C. Margossian, and Julien
Riou. 2021. {``Bayesian Workflow for Disease Transmission Modeling in
{Stan}.''} \emph{Statistics in Medicine} 40 (27): 6209--34.
\url{https://doi.org/10.1002/sim.9164}.

\bibitem[\citeproctext]{ref-grovesBaseball2022}
Groves, Teddy. 2022. {``Baseball.''}
\url{https://github.com/teddygroves/baseball}.

\bibitem[\citeproctext]{ref-grovesteddySphincter2024}
---------. 2024. {``Sphincter.''}
\url{https://github.com/teddygroves/sphincter}.

\bibitem[\citeproctext]{ref-grovesteddyDgfreg2023}
Groves, Teddy, and Jason Jooste. 2023. {``Dgfreg.''} DTU Biosustain.
\url{https://github.com/biosustain/dgfreg}.

\bibitem[\citeproctext]{ref-kumarArviZUnifiedLibrary2019}
Kumar, Ravin, Colin Carroll, Ari Hartikainen, and Osvaldo Martin. 2019.
{``{ArviZ} a Unified Library for Exploratory Analysis of {Bayesian}
Models in {Python}.''} \emph{Journal of Open Source Software} 4 (33):
1143. \url{https://doi.org/10.21105/joss.01143}.

\bibitem[\citeproctext]{ref-niels_bantilan-proc-scipy-2020}
Niels Bantilan. 2020. {``Pandera: {Statistical Data Validation} of
{Pandas Dataframes}.''} In \emph{Proceedings of the 19th {Python} in
{Science Conference}}, edited by Meghann Agarwal, Chris Calloway, Dillon
Niederhut, and David Shupe, 116--24.
\url{https://doi.org/10.25080/Majora-342d178e-010}.

\bibitem[\citeproctext]{ref-pydanticdevelopersPydantic2022}
Pydantic developers. 2022. {``Pydantic.''}
\url{https://pypi.org/project/pydantic/}.

\bibitem[\citeproctext]{ref-standevelopmentteamCmdStanPy2022}
Stan Development Team. 2022. {``{CmdStanPy}.''}
\url{https://github.com/stan-dev/cmdstanpy}.

\bibitem[\citeproctext]{ref-strumbeljPresentFutureSoftware}
Štrumbelj, Erik, Alexandre Bouchard-Côté, Jukka Corander, Andrew Gelman,
Håvard Rue, Lawrence Murray, Henri Pesonen, Martyn Plummer, and Aki
Vehtari. 2024. {``Past, {Present}, and {Future} of {Software} for
{Bayesian Inference}.''}

\end{CSLReferences}

\end{document}