### 2024-02-15 example_projects/baseball/docs/report.qmd
Update baseball example project

---
title: "Baseball"
resource_path: ["../"]
subtitle: "A report created using bibat version 0.3.0"
author: "Teddy Groves"
bibliography: bibliography.bib
reference-location: margin
toc: true
format:
  html:
    embed-resources: true
  pdf: default
filters:
  - include-code-files
---

I used to do a lot of statistical analyses of sports data where there was a
latent parameter for the player's ability. You can see an example
[here](https://github.com/teddygroves/cricket).

It was a natural choice to use a Bayesian hierarchical model where the abilities
have a location/scale type distribution with support on the whole real line, and
in fact this worked pretty well! You would see the kind of players at the top
and bottom of the ability scale that you would expect.

Still, there were a few problems. In particular the data were typically
unbalanced because better players tended to produce more data than worse
players. The result of this was that my models would often inappropriately think
the bad players were like the good players: they would not only tend to be too
certain about the abilities of low-data players, but also be biased, thinking
that these players are probably a bit better than they actually are. I never
came up with a good way to solve this problem, despite trying a lot of things!

Even though I don't work with sports data very much any more, these issues still
lingered in my mind, so when I read [this great case
study](https://mc-stan.org/users/documentation/case-studies/gpareto_functions.html#conclusion-on-the-data-analysis)
about geomagnetic storms it gave me an idea for yet another potential solution.

The idea was this: just as data about intense storms tell us about a tail
of the bigger solar magnetism distribution, maybe data about professional
sportspeople is best thought about as coming from a tail of the general sports
ability distribution. If so, perhaps something like the [generalised pareto
distribution](https://en.wikipedia.org/wiki/Generalized_Pareto_distribution)
might be better than the bog-standard normal distribution for describing the
pros' abilities.

I thought I'd test this out with some sports data, and luckily there is a really
nice baseball example on [the Stan case studies
website](https://mc-stan.org/users/documentation/case-studies/pool-binary-trials.html),
complete with [data from the 2006 Major league
season](https://github.com/stan-dev/example-models/blob/master/knitr/pool-binary-trials/baseball-hits-2006.csv).
After I posted some early results [on the Stan discourse
forum](https://discourse.mc-stan.org/t/baseball-analysis-using-latent-generalised-pareto-distribution/29222),
other users suggested that it might be interesting to model similar data from
different seasons. This data can now be found quite easily on the [baseball
databank](https://baseballdb.lawlesst.net/).

With a few datasets and statistical models to consider, the full analysis looked
like it would be too big to fit in a single file, so this seemed like a job for a
batteries-included Bayesian analysis template like [bibat](https://github.com/teddygroves/bibat).

The rest of this vignette describes how I used bibat to (relatively) painlessly
see if my generalised Pareto distribution idea would work.

Check out [the full analysis](https://github.com/teddygroves/bibat/example_projects/baseball) for all the details.

# Setup

First I installed [copier](https://copier.readthedocs.io/en/stable/) in my
global Python 3.12 environment with this command:

```zsh
> pipx install copier
```

Next I ran bibat's wizard like this:

```{.zsh}
☁  teddygroves  copier copy gh:teddygroves/bibat baseballl
🎤 Name of your project
   Baseball
🎤 Name of your project, with no spaces (used for venv and package names)
   baseball
🎤 A short description of the project.
   Is the generalised Pareto distribution good for modelling latent hitting
ability?
🎤 Your name (or your organization/company/team)
   Teddy Groves
🎤 Author email (will be included in pyproject.toml)
   groves.teddy@gmail.com
🎤 Code of conduct contact
   groves.teddy@gmail.com
🎤 open_source_license
   MIT
🎤 docs_format
   Quarto
🎤 create_dotgithub_directory (bool)
   No
```

After I answered the wizard's questions bibat creted a new folder called
`baseball` that looked like this:

```{.zsh}
☁  teddygroves  tree baseball
baseball
├── CODE_OF_CONDUCT.md
├── LICENSE
├── Makefile
├── README.md
├── data
│   └── raw
│       ├── raw_measurements.csv
│       └── readme.md
├── docs
│   ├── bibliography.bib
│   ├── img
│   │   ├── example.png
│   │   └── readme.md
│   └── report.qmd
├── inferences
│   ├── fake_interaction
│   │   └── config.toml
│   ├── interaction
│   │   └── config.toml
│   └── no_interaction
│       └── config.toml
├── notebooks
│   └── investigate.ipynb
├── plots
│   └── readme.md
├── pyproject.toml
├── src
│   ├── __init__.py
│   ├── data_preparation.py
│   ├── fitting.py
│   ├── stan
│   │   ├── custom_functions.stan
│   │   ├── multilevel-linear-regression.stan
│   │   └── readme.md
│   └── stan_input_functions.py
└── tests
    └── test_integration
        └── test_data_preparation.py
```

This folder implements bibat's example analysis, which compares linear
regression models with different design matrices. To check that everything was
working correctly I ran the analysis like this:

```zsh
$ cd baseball
$ make analysis
```

This ran without errors, running some data preparation functions and then
analysing the data with Stan in a few configurations.

# Getting raw data

To fetch raw data from the internet, I wrote a new script in the file
`baseball/fetch_data.py`:

```{.python include="../src/fetch_data.py"}
```

To get the files I ran the script:

```zsh
(baseball) ☁  baseball  python src/fetch_data.py --verbose
/Users/tedgro/repos/teddygroves/baseball/src/fetch_data.py:6: DeprecationWarning:
Pyarrow will become a required dependency of pandas in the next major release of pandas (pandas 3.0),
(to allow more performant data types, such as the Arrow string type, and better interoperability with other libraries)
but was not found to be installed on your system.
If this would cause problems for you,
please provide us feedback at https://github.com/pandas-dev/pandas/issues/54466

  import pandas as pd
INFO:root:Fetching 2006 data from https://raw.githubusercontent.com/stan-dev/example-models/master/knitr/pool-binary-trials/baseball-hits-2006.csv
INFO:root:Writing 2006 data to data/raw/2006.csv
INFO:root:Fetching bdb-main data from https://raw.githubusercontent.com/cbwinslow/baseballdatabank/master/core/Batting.csv
INFO:root:Writing bdb-main data to data/raw/bdb-main.csv
INFO:root:Fetching bdb-post data from https://raw.githubusercontent.com/cbwinslow/baseballdatabank/master/core/BattingPost.csv
INFO:root:Writing bdb-post data to data/raw/bdb-post.csv
INFO:root:Fetching bdb-apps data from https://raw.githubusercontent.com/cbwinslow/baseballdatabank/master/core/Appearances.csv
INFO:root:Writing bdb-apps data to data/raw/bdb-apps.csv
```

Finally, I removed the example analysis's raw data:

```zsh
> rm data/raw/raw_measurements.csv
```

# Data Preparation

The example project already has some data preparation code in the file
`src/data_preparation.py`: I mostly just had to adapt what was already there.

The first data preparation step was to decide what prepared data should
look like for my analysis and represent this definition using a subclass of
bibat's `PreparedData` [Pydantic `BaseModel`](https://docs.pydantic.dev/latest/api/base_model/). Looking at the example analysis's
`ExamplePreparedData` class, I could see that it already had the components
that I needed, namely a name, a dictionary of coordinates and a table of
measurements:

```{.python include="../../../bibat/template/src/data_preparation.py" start-line=44 end-line=57}
```

There are a few interesting things about this model. First, it uses some
specialised types from bibat's `util` module, namely `CoordDict` and
`DfInPydanticModel`. The latter is quite handy as it ensures that the table of
measurements can be saved to json and then safely read back again. Second, the
measurement table has a validator that refers to a [pandera `DataFrameModel`](https://pandera.readthedocs.io/en/stable/reference/generated/pandera.api.pandas.model.DataFrameModel.html?highlight=SchemaModel) called `ExampleMeasurementsDF` that defines what a measurement table should be like. It is defined just before `ExamplePreparedData`:

```{.python include="../../../bibat/template/src/data_preparation.py" start-line=32 end-line=41}
```

I only had to change a few things to make new definitions for my analysis:

```{.python include="../src/data_preparation.py" start-line=39 end-line=61}
```

The next step is to write functions that return `BaseballPreparedData` objects.
In this case I wrote a couple of data preparation functions: `prepare_data_2006`
and `prepare_data_bdb`:

```{.python include="../src/data_preparation.py" start-line=64 end-line=157}
```

To take into account the inconsistency between my two raw data sources, I first
had to change the variable RAW_DATA_FILES:

```{.python include="../src/data_preparation.py" start-line=29 end-line=36}
```

Next I changed the `prepare_data` function to handle the two different data
sources.

```{.python include="../src/data_preparation.py" start-line=159 end-line=177}
```

To finish off the data preparation step I deleted any unused code from the
example analysis, and updated the function `load_prepared_data`'s signature:

```{.python include="../src/data_preparation.py" start-line=180 end-line=185}
```

To check that all this worked, I ran the data preparation script manually:^[I
could also have just run `make analysis` again. This would have caused an error
on the step after prepare_data.py, which is fine!]

```{.zsh}
> source .venv/bin/activate
(baseball) > python baseball/data_preparation.py
```

Now the folder `data/prepared/` contained json files `2006.json` and `bdb.json`
that looked like this:

```zsh
(baseball) > jq . data/prepared/bdb.json | head -n 10
{
  "name": "bdb",
  "coords": {
    "player_season": [
      "abreujo022017",
      "abreujo022018",
      "abreujo022019",
      "abreujo022020",
      "abreujo022021",
      "acunaro012018",
```

# Specifying statistical models

I wanted to test two statistical models: one with the modelling the distribution
of per-player logit-scale at-bat success rates as a normal distribution with
unknown mean and standard deviation, and another where the same logit-scale
rates have a generalised Pareto distribution.

So, given a table of $N$ player profiles, with each player has $y$ successes out
of $K$ at-bats and an unknown latent success rate $\alpha$, I wanted to use this
measurement model:

$$
y \sim \text{binomial logit}(K, \alpha)
$$

In the generalised Pareto model I would give the $\alpha$s this prior model,
with the hyperparameter $\text{min }\alpha$ assumed to be known exactly and $k$
and $\sigma$ given prior distributions that put the $\alpha$s in the generally
plausible range of between roughly 0.1 and 0.4.

\begin{align*}
\alpha &\sim GPareto(\text{min }\alpha, k, \sigma)
\end{align*}

In the normal model I would use a standard hierarchical regression model with an
effect for the log-scale number of at-bats to attempt to explicitly model the
tendency of players with more appearances to be better:

\begin{align*}
\alpha &\sim Normal(\mu + b_{K} \cdot \ln{K}, \tau) \\
\end{align*}

Again I would choose priors for the hyperparameters that put most of the alphas
between 0.1 and 0.4.

To implement these models using Stan I first added the following function to the
file `custom_functions.stan`. This was simply copied from [the relevant part of
the geomagnetic storms case
study](https://mc-stan.org/users/documentation/case-studies/gpareto_functions.html#conclusion-on-the-data-analysis).

```{.stan}
real gpareto_lpdf(vector y, real ymin, real k, real sigma) {
  // generalised Pareto log pdf
  int N = rows(y);
  real inv_k = inv(k);
  if (k<0 && max(y-ymin)/sigma > -inv_k)
    reject("k<0 and max(y-ymin)/sigma > -1/k; found k, sigma =", k, ", ", sigma);
  if (sigma<=0)
    reject("sigma<=0; found sigma =", sigma);
  if (fabs(k) > 1e-15)
    return -(1+inv_k)*sum(log1p((y-ymin) * (k/sigma))) -N*log(sigma);
  else
    return -sum(y-ymin)/sigma -N*log(sigma); // limit k->0
}
```

Next I wrote a file `gpareto.stan`:

```{.stan include="../src/stan/gpareto.stan"}
```

Finally I wrote a file `normal.stan`:

```{.stan include="../src/stan/normal.stan"}
```

# Generating Stan inputs

Next I needed to tell the analysis how to turn some prepared data into a
dictionary that can be used as input for Stan. The example analysis includes
some functions that do this in the file `src/stan_input_functions.py`.

I followed the examples to create similar functions for my two Stan
models:^[Note that this code uses the scipy function `logit`, which it imported
like this: `from scipy.special import logit`]

```{.python include="../src/stan_input_functions.py" end-line=33}
```

<!-- todo: add link for returns_stan_input below -->

These function use a handy decorator called `returns_stan_input` that uses the
power of [`stanio`](https://github.com/WardBrian/stanio) to convert a dictionary
to a json serialisable, Stan-friendly form. Thanks to this decorator I didn't
have to spend time appending `.values.tolist()` to the inputs that came from
pandas objects.

I now had functions that turned my prepared data into input for either of my
Stan models. But why stop there? It can also be useful to generate fake input
data, by running a model in simulation mode using hardcoded parameter values.
Here are some functions that do this for both of my models: ^[These functions
require some more imports: `numpy` and `scipy.special.expit`]

```{.python include="../src/stan_input_functions.py" start-line=37}
```

# Configuring fitting

To tell my analysis how to run inferences, I had to edit the file `src/
fitting.py`. The only changes I made from the example analysis were to remove
the `kfold` fitting mode, which I didn't need to use in this analysis, and to
add my Stan input functions to the `LOCAL_FUNCTIONS` constant:

```{.python include="../src/fitting.py"}
```

# Specifying inferences

Now all the building blocks for making statistical inferences – raw data, data
preparation rules, statistical models, recipes for turning prepared data into
model inputs and a procedure for running inferences – were in place. The last
step before actually running my analysis was to write down how put the blocks
together. Bibat has another concept for this, called 'inferences'.

An inference in bibat is a folder containing a special file called
`config.toml`. This file sets out what inferences you want to make: which
statistical model, which prepared data function, which Stan input function,
which parameters have which dimensions, which sampling modes to use and how to
configure the sampler. The folder will later be filled up with the results of
performing the specified inferences.

I started by deleting the example inferences and creating two fresh folders,
leaving me with an `inferences` folder looking like this:

```zsh
inferences
├── gpareto2006
│   └── config.toml
└── normal2006
    └── config.toml
```

Here is the file `inferences/gpareto2006/config.toml`:

```{.toml include="../inferences/gpareto2006/config.toml"}
```

Here is the file `inferences/normal2006/config.toml`:

```{.toml include="../inferences/normal2006/config.toml"}
```

Note that:

- The Stan file, prepared data folder and Stan input function are referred to by
  strings. The analysis should raise an error if you enter a non-existing value.
- Both inferences are set to run in `prior` and `posterior` modes
- You can enter arbitrary arguments to cmdstanpy's `CmdStanModel.sample` method
  in the `[sample_kwargs]` table.
- You can enter mode-specific overrides in `[mode_options.<MODE>]`. This can be
  handy if you want to run more or fewer iterations for a certain mode.

With all these changes made I ran `make analysis` again. There were no errors
until after sampling, which was expected as I hadn't yet customised the
investigation code, and I saw messages indicating that Stan had run. I also
found that the `inferences` subfolders had been populated:

```zsh
inferences
├── gpareto2006
│   ├── config.toml
│   └── idata
└── normal2006
    ├── config.toml
    └── idata
```

# Investigating the inferences

Now that the inferences are ready it's time to check them out. Bibat provides a
Jupyter notebook at `notebooks/investigate.ipynb` for exactly this purpose. The
notebook's main job is to create plots and save them in the `plots` directory
when it is executed with the command `jupyter execute notebooks/investigate.ipynb`,
 which is the final step in the chain of commands that is triggered by `make
analysis`.

A notebook is arguably a nicer home for code that creates plots than a plain
python script because it allows for literate documentation and an iterative
workflow. A notebook makes it easy to, for example, add some code to change the
scale of a plot, execute the code and see the new results, then update the
relevant documentation all in the same place.

The code from the example analysis's notebook for loading `InferenceData` was
reusable with a few tweaks to avoid missing file errors, so I kept it. On the
other hand, I wanted to make some different plots from the ones in the example
analysis, including some that required loading prepared data. To check out
everything I did, see
[here](https://github.com/teddygroves/bibat/blob/main/bibat/example_projects/baseball/notebooks/investigate.ipynb).

# Choosing priors using push-forward calibration

The trickiest thing about my analysis was setting prior distributions for the
parameters $k$ and $\sigma$ in the generalised Pareto models. To choose some
more or less plausible values I did a few prior-mode model runs and checked the
distributions of the resulting `alpha` variables. I wanted to make sure that
they all lay in the range corresponding to batting averages between about 0.1
and a little over 0.4. Here is a graph that shows the 1% to 99% prior quantiles
for each player's latent success percentage in both datasets alongside their
actually realised success rates.

![](../plots/prior_quantiles.png)

# Extending the analysis to the baseballdatabank data

To model the more recent data, all I had to do was create some new inference
folders with appropriate `prepared_data` fields in their `config.toml`
files. For example, here is the `config.toml` file for the `gparetobdb`
inference:

```{.toml include="../inferences/gparetobdb/config.toml"}
```

After running `make analysis` one more time, I went back to the notebook
`notebooks/investigate.ipynb` and made plots of both models' posterior 1%-99%
success rate intervals for both datasets:

![](../plots/posterior_quantiles.png)

I think this is very interesting. Both models' prior distributions had similar
regularisation levels, and they more or less agree about the abilities of the
players with the most at-bats, both in terms of locations and the widths of the
plausible intervals for the true success rate. However, the models ended up with
dramatically different certainty levels about the abilities of players with few
at-bats. This pattern was true both for the small 2006 dataset and the much
larger baseballdatabank dataset.

# Documenting the analysis

The final step was to document my analysis. To do this I edited the file
`docs/report.qmd`, then ran `make docs`, which produced the
very HTML document that you are probably reading now! You can find the complete
`report.qmd` file
[here](https://github.com/teddygroves/bibat/blob/main/example_projects/baseball/docs/report.qmd).