### 2024-02-26 paper/automl-template.tex
add automl paper stuff

\documentclass[11pt]{article}

\usepackage{microtype} % microtypography
\usepackage{booktabs}  % tables
\usepackage{url}  % urls

% With no package options, the submission will be anonymized, the supplemental
% material will be visible, and line numbers will be added to the manuscript.
%
% To compile a non-anonymized camera-ready version, add the [final] option:
%
% \usepackage[final]{automl}
%
% To compile for the workshop track, add the [workshop] option:
%
% \usepackage[workshop]{automl}
% \usepackage[final,workshop]{automl}
%
% To compile a single-blind submission, e.g. for the ABCD track, add the
% [revealauthors] option:
%
% \usepackage[revealauthors]{automl}
%
% To compile a non-anonymized pre-print version that can be uploaded to arXiv,
% add the [preprint] option:
%
% \usepackage[preprint]{automl}
%
% The [hidesupplement] option may be used to hide the supplementary material from any of the
% above commands, e.g.:
%
% \usepackage[hidesupplement,final,workshop]{automl}

\usepackage[revealauthors]{automl}

$pandoc.tex()$
$title.tex()$

\begin{document}
\maketitle
\begin{abstract}
$abstract$
\end{abstract}
$body$

\newpage
\section*{Submission Checklist}

% All submissions must include a section containing the AutoML submission
% checklist, shown below. The submission checklist does not count towards the
% page limit and must be included at submission and camera-ready time.  The
% submission checklist draws upon related submission checklists:
%
% the NeurIPS '22 checklist: https://neurips.cc/Conferences/2022/PaperInformation/PaperChecklist
%     and the NAS checklist: https://www.automl.org/wp-content/uploads/NAS/NAS_checklist.pdf
%
% For each question, change the default \answerTODO{} to either:
%
%     \answerYes{[justification]},
%     \answerNo{[justification]}, or
%     \answerNA{[justification]}.

% *You must include a brief justification to your answer,} either by referencing
% *the appropriate section of your paper or providing a brief inline
% *description.* For example:
%
% -  Did you include the license of the code and datasets?
%
%    \answerYes{See Section~\ref{sec:code}.
%
% - Did you include all the code for running experiments?
%
%   \answerNo{We include the code we wrote for conducting the experiments, but
%   complete replication depends on proprietary libraries for executing on a
%   private compute cluster. The code therefore is not runnable without
%   modification. To compensate, we provide a runnable but non-parallelized
%   version of the code that could replicate the results at the expense of a
%   greater wall-clock time.}
%
% - Did you include the license of the datasets?
%
%   \answerNA{Our experiments were conducted on publicly available datasets and
%   we have not introduced new datasets.}
%
% Please note that if you answer a question with \verb|\answerNo{}|, we expect
% that you provide an explanation and/or compensation for the omission. For
% example, if you cannot provide complete evaluation code for some reason, you
% might instead provide code for a minimal reproduction of the main insights of
% your paper.
%
% Please do not modify the questions and only use the provided macros for your
% answers. Note that the submission checklist does not count towards the page
% limit.

\begin{enumerate}
\item For all authors\dots
  %
  \begin{enumerate}
  \item Do the main claims made in the abstract and introduction accurately
    reflect the paper's contributions and scope?
    %
    \answerYes{}
    %
  \item Did you describe the limitations of your work?
    %
    \answerYes{See section ``Limitations''}
    %
  \item Did you discuss any potential negative societal impacts of your work?
    %
    \answerYes{}
    %
  \item Did you read the ethics review guidelines and ensure that your paper
    conforms to them? \url{https://2022.automl.cc/ethics-accessibility/}
    %
    \answerYes{}
    %
  \end{enumerate}
  %
\item If you ran experiments\dots
  %
  \begin{enumerate}
  \item Did you use the same evaluation protocol for all methods being compared (e.g.,
    same benchmarks, data (sub)sets, available resources)?
    %
    \answerNA{}
    %
  \item Did you specify all the necessary details of your evaluation (e.g., data splits,
    pre-processing, search spaces, hyperparameter tuning)?
    %
    \answerNA{}
    %
  \item Did you repeat your experiments (e.g., across multiple random seeds or splits) to account for the impact of randomness in your methods or data?
    %
    \answerNA{}
    %
  \item Did you report the uncertainty of your results (e.g., the variance across random seeds or splits)?
    %
    \answerNA{}
    %
  \item Did you report the statistical significance of your results?
    %
    \answerNA{}
    %
  \item Did you use tabular or surrogate benchmarks for in-depth evaluations?
    %
    \answerNA{}
    %
  \item Did you compare performance over time and describe how you selected the maximum duration?
    %
    \answerNA{}
    %
  \item Did you include the total amount of compute and the type of resources
    used (e.g., type of \textsc{gpu}s, internal cluster, or cloud provider)?
    %
    \answerNA{}
    %
  \item Did you run ablation studies to assess the impact of different
    components of your approach?
    %
    \answerNA{}
    %
  \end{enumerate}
  %
\item With respect to the code used to obtain your results\dots
  %
  \begin{enumerate}
\item Did you include the code, data, and instructions needed to reproduce the
    main experimental results, including all requirements (e.g.,
    \texttt{requirements.txt} with explicit versions), random seeds, an instructive
    \texttt{README} with installation, and execution commands (either in the
    supplemental material or as a \textsc{url})?
    %
    \answerYes{}
    %
  \item Did you include a minimal example to replicate results on a small subset
    of the experiments or on toy data?
    %
    \answerYes{}
    %
  \item Did you ensure sufficient code quality and documentation so that someone else
    can execute and understand your code?
    %
    \answerYes{See pyopensci review \url{https://github.com/pyOpenSci/software-submission/issues/83}}
    %
  \item Did you include the raw results of running your experiments with the given
    code, data, and instructions?
    %
    \answerNA{}
    %
  \item Did you include the code, additional data, and instructions needed to generate
    the figures and tables in your paper based on the raw results?
    %
    \answerYes{}
    %
  \end{enumerate}
  %
\item If you used existing assets (e.g., code, data, models)\dots
  %
  \begin{enumerate}
  \item Did you cite the creators of used assets?
    %
    \answerYes{}
    %
  \item Did you discuss whether and how consent was obtained from people whose
    data you're using/curating if the license requires it?
    %
    \answerNA{}
    %
  \item Did you discuss whether the data you are using/curating contains
    personally identifiable information or offensive content?
    %
    \answerNA{}
    %
  \end{enumerate}
  %
\item If you created/released new assets (e.g., code, data, models)\dots
  %
  \begin{enumerate}
    %
    \item Did you mention the license of the new assets (e.g., as part of your code submission)?
    %
    \answerYes{}
    %
    \item Did you include the new assets either in the supplemental material or as
    a \textsc{url} (to, e.g., GitHub or Hugging Face)?
    %
    \answerYes{}
    %
  \end{enumerate}
  %
\item If you used crowdsourcing or conducted research with human subjects\dots
  %
  \begin{enumerate}
  \item Did you include the full text of instructions given to participants and
    screenshots, if applicable?
    %
    \answerNA{}
    %
  \item Did you describe any potential participant risks, with links to
    Institutional Review Board (\textsc{irb}) approvals, if applicable?
    %
    \answerNA{}
    %
  \item Did you include the estimated hourly wage paid to participants and the
    total amount spent on participant compensation?
    %
    \answerNA{}
    %
  \end{enumerate}
\item If you included theoretical results\dots
  %
  \begin{enumerate}
  \item Did you state the full set of assumptions of all theoretical results?
    %
    \answerNA{}
    %
  \item Did you include complete proofs of all theoretical results?
    %
    \answerNA{}
    %
  \end{enumerate}
  %
\end{enumerate}

\end{document}