### 2024-02-26 paper/main.qmd
Update paper

---
title: "Bibat: Batteries-include Bayesian Analysis Template"
abstract: >
  Bayesian statistical workflow offers a powerful way to learn from data,
  but software software projects that implement complex Bayesian workflows
  in practice are unusual, partly due to the difficulty of orchestrating
  Bayesian statistical software. Bibat addresses this challenge by providing
  a full-featured, scalable Bayesian statistical analysis project using
  an interactive template. Bibat is available on the Python Package index,
  documented at <https://bibat.readthedocs.io/> and developed at <https://github.com/teddygroves/bibat/>. Bibat is free to use under the MIT license. This paper explains the motivation for bibat, describes intended usage, discusses bibat's design, compares bibat with similar software, highlights several examples of bibat's use in science and provides links to community resources associated with bibat.
number-sections: true
author:
  - name: "Teddy Groves"
    orcid: 0000-0002-7109-3270
    email: tedgro@biosustain.dtu.dk
    affiliations:
      - id: 1
        name: Danish Technical University
        department: The Novo Nordisk Foundation Center for Biosustainability
        city: Kongens Lyngby
        country: Denmark
    attributes:
      corresponding: true
bibliography: bibliography.bib
keywords: [Bayesian workflow, Methodology, Software]
format:
  html:
    embed-resources: true
  pdf:
    template: automl-template.tex
    template-partials:
      - title.tex
---

# Introduction: the problem of orchestrating Bayesian workflow software

The term "Bayesian workflow" captures the idea that Bayesian statistical
analysis comprises not just inference, but also specific approaches to related
activities like data preparation, model design, diagnosis, debugging and
criticism. This idea can be found in @boxBayesianInferenceStatistical1992
and has recently received increasing scholarly recognition
[@gelmanBayesianWorkflow2020;@grinsztajnBayesianWorkflowDisease2021;
@gabryVisualizationBayesianWorkflow2019]. Software tools now exist that address
most individual aspects of a Bayesian workflow: see @vstrumbelj2024past for a
review of the state of the art.

Unfortunately, each tool typically addresses one, or at most a few, of the many
activities that comprise a real Bayesian workflow software project; it is left
to the individual project team to orchestrate all of the tools they require.
Writing software that performs this orchestration can be time-consuming and
tricky, especially in the common scenario where it is not initially clear how
many, or what kind of, statistical models, datasets, data manipulations or
investigations an analysis will require.

Bibat is a new tool that addresses the difficulty of orchestrating Bayesian
workflow software by providing a full-featured, high-quality project that can be
extended to implement a wide range of statistical analyses.

# How bibat works

In order to use bibat, a user must first install the templating library copier
[@copierdevelopersCopier2024], then choose a directory name and run this command
from the command line:

```sh
$ copier copy gh:teddygroves/bibat my-chosen-directory-name
```

This command triggers an interactive form which prompts the user to select
a range of customisation options. The new directory will then be created
if necessary and filled with code that implements an example analysis, with
customisations reflecting the user's choices. This analysis works immediately,
and can be reproduced with the single command `make analysis` without the need
for any further action by the user: in this sense bibat comes with batteries
included.

@fig-workflow illustrates the components of a bibat-based Bayesian workflow and
shows how it proceeds: the project team edits the code components, then runs
`make analysis`, triggering creation of the result components. After inspecting
these they repeat the process, leading to a cycle that ultimately results in a
complete, easily reproducible analysis.

:::{#fig-workflow}

![](img/workflow.png){width=100%}

Schematic representation of a Bayesian workflow implemented using bibat. The
author inspects their analysis's results, edits code corresponding to the
boxes on the left, runs the command `make analysis`, then repeats. The diagram
illustrates several key features of bibat: inference components are modular and
plural, the overall workflow is iterative and cyclical and the whole analysis
can be executed with a single command.

:::

Bibat is documented at <https://github.com/teddygroves/bibat/>. The
documentation website includes instructions for getting started, a detailed
explanation of bibat's concepts and an extended vignette illustrating how to
implement a complex statistical analysis starting from bibat's example analysis
usage. In addition, the documentation site contains a full description of
bibat's python API and command line interface, instructions for contributing and
a section discussing accessibility considerations.

# Design choices

Bibat's design was informed by the aims to accommodate the many sources of
complexity and in a Bayesian workflow project, to ensure easy reproducibility
and to integrate many open-source, widely-adopted and powerful Bayesian workflow
tools and to encourage collaborative development.

As discussed in @gelmanBayesianWorkflow2020, Bayesian workflows are complicated,
featuring plurality, cyclicity and complexity at many levels. As a specialised
Bayesian workflow template, a key goal for bibat was to manage this complexity.
Bibat achieves this aim by separating non-interacting analysis components into
separate, potentially plural modules and by serialising data to files wherever
possible. Prepared datasets, statistical models, inference configurations,
inference results, plots and analyses all have file representations. Fitting
modes, data manipulations and data models are modularised in code through the
use of appropriately structured data classes and functions. Thanks to this
modular approach it is possible to perform small sub analyses individually
and to iteratively expand the analysis by adding components without needing to
consider everything at once. In addition, bibat ensures that there are minimal
restrictions on the components: for example, datasets need not be singular or
tabular, and it is possible to use any statistical model that Stan can compile.
Thanks to these accommodations a project team using bibat should typically not
need to foresee the ultimate requirements of their analysis before starting
the project.

Bibat encourages reproducibility by providing a preconfigured makefile
with a target `analysis` that triggers creation of an isolated environment,
installation of dependencies, data preparation, statistical computation and
analysis of results. In this way a bibat analysis can be reproduced on most
platforms using a single command. In particular, this target attempts to install
cmdstan if necessary, using a recipe tailored to the host operating system.
This functionality addresses a common issue where researchers find it difficult
to install Stan, especially on Windows. A second way in which bibat encourages
reproducibility is by providing a preconfigured Python project following modern
conventions, making bibat analyses straightforward to replicated and extend for
other researchers who are already familiar with these conventions.

Bibat integrates many widely-adopted open-source tools to implement
the components of a Bayesian workflow. These include pydantic
[@pydanticdevelopersPydantic2022] and pandera [@niels_bantilan-proc-scipy-2020]
for data modelling, Stan [@carpenterStanProbabilisticProgramming2017] for
statistical inference, cmdstanpy [@standevelopmentteamCmdStanPy2022] for
python-Stan interface, arviz [@kumarArviZUnifiedLibrary2019] for storing and
analysing inferences and sphinx [@georgbrandlandthesphinxteamSphinx2022] and
quarto [@Allaire_Quarto_2022] for documentation.

To encourage collaborative development of Bayesian workflow projects, Bibat
projects include a preconfigured test environment, continuous integration,
linting and pre-commit hooks. In addition, bibat includes documentation as a
first class, integrated component of the analysis, helping to keep it in sync
with the other components.

Bibat is continuously tested to ensure that it works on the operating systems
Linux, macOS and Windows. Bibat's continuous integration runs a test suite as
well as an end-to-end functional test on all supported Python versions.

# Fitting modes

The most novel part of bibat's design is the introduction of an abstraction
called "fitting mode", which allows bibat projects to handle fitting a model
to a dataset in different ways. This is often necessary as part of a Bayesian
workflow: for example, one might perform MCMC sampling of both the prior
and posterior distributions, or perform multiple leave-out-one-fold fits for
cross-validation, or compare MCMC-based posterior inference with an alternative
computation method.

Fitting modes in bibat projects take the form of instances of the class
`FittingMode`. Each fitting mode contains a name, a function that fits a
prepared dataset and instructions for how and where to save the results. For
example, the provided prior sampling fitting mode is called "prior", contains
a function that runs MCMC sampling with the `likelihood` input variable set to
`0`, returning a `CmdStanMCMC` instance, and specifies that this result should
be written to the `InferenceData` group `prior`.  Bibat provides fitting modes
corresponding for prior sampling, posterior sampling and k-fold posterior
sampling. Users can easily implement additional fitting modes or modify the
`FittingMode` class to achieve even richer functionality. Fitting modes can be
referenced by name from the file that configures an inferece: for example, the
following lines indicate that the inference should be run in prior, posterior
and kfold modes:

```toml
modes = ["prior", "posterior", "kfold"]
```

Fitting modes allow bibat projects to succinctly but flexibly declare how to
perform inferences, and allow results corresponding to the same inference to be
stored alongside each other appropriately.

# Comparison with alternative software

Other than bibat, we are not aware of any interactive template that specifically
targets Bayesian workflow projects. There are some templates that arguably
encompass Bayesian workflow as a special case of data analysis project,
such as cookiecutter-data-science [@drivendataCookiecutterdatascience2022],
but these are of limited use compared with a specialised template due to
the many specificities of Bayesian workflow. cookiecutter-cmdstanpy-wrapper
[@wardWardBrianCookiecuttercmdstanpywrapper2024] is an interactive template that
targets a different use case than Bayesian workflow projects, namely setting up
a Python package that provides pre-compiled Stan models.

There is some software that addresses the general task of facilitating
Bayesian workflow, but using a different approach from bibat's. For example,
bambi [@capretto2020] and brms [@burknerBrmsPackageBayesian2017] aim to make
implementing Bayesian workflows easier by providing ergonomic ways to specify
and fit Bayesian regression models to tabular datasets. Bibat is complementary
with these packages, as it targets use cases that they do not support, such as
analyses where complex datasets or custom models might be required.

# Limitations

Using bibat effectively requires familiarity with Pydantic, pandera, arviz, Stan
and managing a medium-sized Python project. Many statistical analysis projects
do not require using these tools, for example if data preparation or validation
is trivial, if custom statistical models are not required, or if the analysis
can be carried out by a single script. Practitioners who wish to implement such
Bayesian workflows may prefer to simply write their software from scratch rather
than use bibat, using tools like bambi or brms to ensure that the software
challenge remains manageable.

Similarly, bibat accommodates plural inferences, fitting modes and datasets,
but many analyses are singular in at least one of these components and could
therefore be implemented more simply and concisely than an equivalent bibat
project. On the other hand, it is typically difficult to predict in advance
which components of a Bayesian workflow will be plural, and costly to re-write
a project after mistakenly assuming that a component will be singular. While we
acknowledge that accommodating potentially unneeded plurality is an important
limitation of bibat, we nonetheless think that it is the correct choice for a
general-purpose template.

Another limitation of bibat is that it makes many opinionated choices about
which tools to use. In particular, languages other than Python, inference
frameworks other than Stan and validation frameworks other than Pydantic
are not supported. We think that it is on the whole good for templates to be
opinionated, as unopinionated templates are necessarily more complicated; this
limitation of bibat is therefore best addressed by the development of additional
analysis templates that make different choices.

# Case studies

The following cases illustrate how bibat has been used in practice to facilitate
Bayesian workflow projects.

@grovesteddyDgfreg2023 used bibat to compare a Bayesian and two non-Bayesian
approaches to modelling a biochemical thermodynamics dataset. Bibat facilitated
this analysis even though it was not very large---the final analysis contained
one dataset, three models and three inferences---because the fitting mode
abstraction allowed for straightforward comparison of the different methods.
Additionally, bibat made it easier to iteratively investigate and discard models
that did not form part of the final analysis.

In @grovesBaseball2022, Bibat was used to implement a sports analysis involving
two datasets, two models and four inferences, demonstrating that the generalised
Pareto distribution can be used to describe hitting ability in baseball. This
analysis is now included in bibat as an illustration, along with an accompanying
tutorial. An illustrative graphic from this analysis is shown in @fig-baseball.

In this case bibat was useful because of its ability to implement arbitrary
statistical models, as latent generalised Pareto distributions are not supported
by any available formula-based regression packages. Further, bibat's modular
design made it easier to implement this medium-sized analysis with two datasets,
two models and six inferences.

:::{#fig-baseball}

![](img/baseball.png){width=100%}

A graphical posterior predictive check produced as part of a bibat
analysis that fit two statistical models to two datasets of baseball data.
The coloured lines show each model's posterior predictive distributions
and the black dots show the two observed datasets.  See <https://github.com/teddygroves/bibat/tree/main/bibat/examples/baseball>
for the full analysis.

:::

These cases illustrate that bibat can be useful in a variety of real Bayesian
workflows, with different sizes, subject matters and emphases.

# Community

Bibat is developed in public and encourages community contribution. Please see the contributing page <https://github.com/teddygroves/bibat/blob/main/CONTRIBUTING.md> and code of conduct <https://github.com/teddygroves/bibat/blob/main/CODE_OF_CONDUCT.md> if you would like to help develop bibat.

Bibat has a growing user community, with 16 GitHub stars at the time of
writing, and is affiliated with cmdstanpy through a link on its documentation
website. Bibat is also affiliated with the Python scientific software community
PyOpenSci, allowing for help with maintenance as well as peer review for
code and documentation quality, usability and accessibility. The PyOpenSci
peer review for bibat can be found here: <https://github.com/pyOpenSci/software-submission/issues/83>.

# Broader impact statement

After careful reflection, the authors have determined that this work presents no
notable negative impacts to society or the environment.

\begin{acknowledgements}

This research was funded by the Novo Nordisk Foundation (Grant numbers
NNF20CC0035580 NNF14OC0009473).

\end{acknowledgements}